---
title: "ST303 Group6 Project"
output: html_document
date: "2025-04-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
We compute the exact values of C(t) for t values from 1 to 100.
```{r include=FALSE}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
```

### Monte Carlo Methods
We take n=10^6.
```{r include=FALSE}
n <- 10^6
```

#### Transformation 1

For transformation 1, we use $u=sinx$, then $x=arcsin(u)$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{x=\arcsin(u)}{=} \int_{0}^{1} {u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[{u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}}\right] \overset{MC}{\approx} \sum_{i=1}^{N}[u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}]
$$

##### Using for loop to investigate the performance of integral across different values of t:
```{r include=FALSE}
set.seed(1)
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_1 <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample_1)
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample_1)/sqrt(n)
}

rbind(mean_1=mean(sample_1),mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2
For transformation 2, we use $u=cosx$, then $x=\arccos(u)$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{ x=\arccos(u)}{=} \int_{0}^{1} {u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[{u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}}\right] \overset{MC}{\approx} \sum_{i=1}^{N}[u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}]
$$

##### Using for loop to investigate the performance of integral across different values of t:
```{r include=FALSE}
set.seed(1)
EstIntegral_2 <- numeric(length(grid_t))
Diff_2 <- numeric(length(grid_t))
Std_Error_2 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_2 <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_2[t] <- mean(sample_2)
  Diff_2[t] <- exact_formula_C(t) - EstIntegral_2[t]
  Std_Error_2[t] <- sd(sample_2)/sqrt(n)
}

rbind(mean_2=mean(sample_2),mean_diff_2 = mean(Diff_2), mean_SE_2 = mean(Std_Error_2))
```
According to the above results, transformation 1 and 2 have exactly the same mean and standard error.

#### Transformation 3
For transformation 3, we use $u=\frac{2}\pi*x$, then $x=\frac{\pi}2u$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:

$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{x=\frac{\pi}2u}{=} \int_{0}^{1} {[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]\frac{\pi}2}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[\frac{\pi}2[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]\right] \overset{MC}{\approx} \sum_{i=1}^{N}\frac{\pi}2[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]
$$

##### Using for loop to investigate the performance of integral across different values of t:
```{r include=FALSE}
set.seed(1)
EstIntegral_3 <- numeric(length(grid_t))
Diff_3 <- numeric(length(grid_t))
Std_Error_3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_3 <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_3[t] <- mean(sample_3)
  Diff_3[t] <- exact_formula_C(t) - EstIntegral_3[t]
  Std_Error_3[t] <- sd(sample_3)/sqrt(n)
}

rbind(mean_diff_3 = mean(Diff_3), SE_3 = mean(Std_Error_3))
```

The following table shows the comparison between different methods:
```{r echo=FALSE}
df_ts<-data.frame(
  Transformation = c("Transformation 1","Transformation 2","Transformation 3"),
  Mean_Diff=c(mean(Diff_1),mean(Diff_2),mean(Diff_3)),
  Mean_Standard_Error=c(mean(Std_Error_1),mean(Std_Error_2),mean(Std_Error_3))
)
df_ts
```

According to the transformation methods above, transformation 3 has the smallest difference between estimated mean and actual mean, and also the smallest standard error. Thus we do the variance reduction based on transformation 3.

The antithetic variate reduction is not applicable since the function is not monotone. Then we try control variates method first.

#### 1.Control variates method

We want to find a random variable $Y$ with mean $0$ and a good constant $c$ such that $\mathbb{E}(X+cY) \approx \theta$.

\begin{itemize}
  \item Simulate $U_1, U_2, \ldots, U_n \overset{d}{\sim} \mathcal{U}(0,1)$.
  \item Simulate $\tilde{X}=(\frac{\pi}{2}u-\frac{(\frac{\pi}{2}u)^{3}}{6})^{t}+(1-\frac{(\frac{\pi}{2}u)^{2}}{2})^{t}$, where $(\frac{\pi}{2}u-\frac{(\frac{\pi}{2}u)^{3}}{6})$ is the Taylor expansion of $\sin(\frac{\pi}{2}u)$, and $(1-\frac{(\frac{\pi}{2}u)^{2}}{2})$ is the Taylor expansion of $\cos(\frac{\pi}{2}u)$.
  \item Simulate $Y=\tilde{X}-\mathbb{E}\left(\tilde{X}\right)$.
  \item Calculate the $\text{Cov}(X,Y)$ and $\text{Var}(Y)$ with the samples to determine $c^{\star}$, $c^{\star}=-\frac{\text{Cov}(X,Y)}{\text{Var}(Y)}$
  \item Approximate $\theta$ with the mean of $X+c^{\star}Y$ and standard errors with the standard errors of $X+c^{\star}Y$.
\end{itemize}

###### Control Variate Method
```{r include=FALSE}
set.seed(1)
EstIntegral_cv3 <- numeric(length(grid_t))
Diff_cv3 <- numeric(length(grid_t))
SE_cv3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  X <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  X_tilde <- ((pi/2*u - (pi/2*u)^3/6)^t + (1 - (pi/2*u)^2/2)^t) * pi/2
  Y <- X_tilde - mean(X_tilde)
  
  c <- -cov(X,Y)/var(Y)
  X_cv3 <- X+c*Y
  EstIntegral_cv3[t] <- mean(X_cv3)
  Diff_cv3[t] <- exact_values_C[t] -  mean(X_cv3)
  SE_cv3[t] <- sd(X_cv3)/sqrt(n)
}
```

#### Stratified Sampling
We estimate:
\begin{align*}
h(U)=[(\sin(\frac{\pi}{2}\frac{U_j + j - 1}{n}))^{t} + (\cos(\frac{\pi}{2}\frac{U_j + j - 1}{n}))^t]\frac{\pi}{2}
\end{align*}

```{r include=FALSE}
set.seed(1)

EstIntegral_ss <- numeric(length(grid_t))
Diff_ss <- numeric(length(grid_t))
Std_Error_ss <- numeric(length(grid_t))

for (t in grid_t){
  j <- 1:n  # split the integral [0, 1] into 10^5 smaller intervals
  u <- runif(n)
  
  sample_ss <- (sin(pi/2*(u + j - 1)/n)^t + cos(pi/2*((u + j - 1)/n))^t) * pi/2
  EstIntegral_ss[t] <- mean(sample_ss)
  Diff_ss[t] <- exact_formula_C(t) - EstIntegral_ss[t]
  Std_Error_ss[t] <- sd(sample_ss)/sqrt(n)
}
```

#### Stratified Sampling + Control Variates

$$
X=[(\sin(\frac{\pi}{2}\frac{U_j + j - 1}{n}))^{t} + (\cos(\frac{\pi}{2}\frac{U_j + j - 1}{n}))^t]\frac{\pi}{2}
$$

$$
\tilde{X}=[(\frac{\pi}{2}\frac{U_j + j - 1}{n}-\frac{\pi}{2}\frac{(\frac{U_j+j-1}{n})^{3}}{6})^{t} + (1-\frac{(\frac{\pi}{2}\frac{U_j+j-1}{n})^{2}}{2})^t]\frac{\pi}{2}
$$
$$
Y=\tilde{X}-\mathbb{E}\left(\tilde{X}\right)
$$

Calculate the $\text{Cov}(X,Y)$ and $\text{Var}(Y)$ with the samples to determine $c^{\star}$, $c^{\star}=-\frac{\text{Cov}(X,Y)}{\text{Var}(Y)}$. \
Approximate $\theta$ with the mean of $X+c^{\star}Y$ and standard errors with the standard errors of $X+c^{\star}Y$.

```{r include=FALSE}
EstIntegral_cvss <- numeric(length(grid_t))
Diff_cvss <- numeric(length(grid_t))
SE_cvss <- numeric(length(grid_t))

for (t in grid_t){
  j <- 1:n  # split the integral [0, 1] into 10^5 smaller intervals
  u <- runif(n)

  X <- (sin(pi/2*(u + j - 1)/n)^t + cos(pi/2*(u + j - 1)/n)^t) * pi/2
  X_tilde <- ((pi/2*(u + j - 1)/n - (pi/2*(u + j - 1)/n)^3/6)^t + (1 - (pi/2*(u + j - 1)/n)^2/2)^t) * pi/2
  Y <- X_tilde - mean(X_tilde)
  
  c <- -cov(X,Y)/var(Y)
  X_cvss <- X+c*Y
  EstIntegral_cvss[t] <- mean(X_cvss)
  Diff_cvss[t] <- exact_values_C[t] -  mean(X_cvss)
  SE_cvss[t] <- sd(X_cvss)/sqrt(n)
}
```

#### Compare all variance reduction methods
```{r}
# Compare SS + CV with pure Monte Carlo without any variance reduction
(abs(Diff_3) - abs(Diff_cvss))[1:10]
(round((Std_Error_3 - SE_cvss)/Std_Error_3, digits = 2))[1:10]

# Compare CV with pure Monte Carlo without any variance reduction
(abs(Diff_3) - abs(Diff_cv3))[1:10]
(round((Std_Error_3 - SE_cv3)/Std_Error_3, digits = 2))[1:10]
```

This table represents the differences between two variance reduction methods (SS+CV, only CV) with Monte Carlo Methods for $t=1,2,...,10$.

```{r}
compare_df <- data.frame(
  t = 1:10,
  diff_SSCV = (abs(Diff_3) - abs(Diff_cvss))[1:10],
  sd_improved_SSCV = (round((Std_Error_3 - SE_cvss)/Std_Error_3, digits = 2))[1:10],
  diff_CV = (abs(Diff_3) - abs(Diff_cv3))[1:10],
  sd_improved_CV = (round((Std_Error_3 - SE_cv3)/Std_Error_3, digits = 2))[1:10]
)
compare_df
```

## (b)
### Calculate the true value of the double integral
```{r echo=TRUE}
set.seed(1)
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
true_integral_1b <- result$value
true_integral_1b
```

### Pure Monte Carlo Methods
Since the equation is symmetric, we can compute: 
$$
    \int_{-1}^{1}\int_{-1}^{1}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy{=}4\int_{0}^{1}\int_{0}^{1}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy  
    \overset{(X, Y) ind \overset{d}{\sim} \mathcal{U}(0,1) }{=}\\
    4\mathbb{E}[\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}} ] \overset{MC}{\approx} \frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy ,
$$

where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.

```{r include=FALSE}
set.seed(1)

n <- 10^5
x <- runif(n)
y <- runif(n)
sample_mc <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)

EstIntegral_mc <- mean(sample_mc)
Diff_mc <- true_integral_1b - EstIntegral_mc
Std_Error_mc <- sd(sample_mc)/sqrt(n)

rbind(EstIntegral_mc, Diff_mc, Std_Error_mc)
```

#### 1.Antithetic Variate Reduction
We need to calculate
$$
\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(x^2 + y^2)})}{\sqrt{x^2 + y^2}}=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{((1-x)^2 + (1-y)^2)})}{\sqrt{(1-x)^2 + (1-y)^2}},
$$

where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.
```{r include=FALSE}
set.seed(1)
n <- 10^5
x <- runif(n)
y <- runif(n)
sample1 <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)
sample2 <- 4 * sin(sqrt((1-x)^2 + (1-y)^2))/sqrt((1-x)^2 + (1-y)^2)
EstIntegral_av <- mean((sample1+sample2)/2)
Mean_diff_av <- true_integral_1b - EstIntegral_av
Std_Error_av <- sd((sample1+sample2)/2)/sqrt(n)

rbind(EstIntegral_av, Mean_diff_av, Std_Error_av)
```

```{r echo=TRUE}
improve_av <- (Std_Error_mc - Std_Error_av)/Std_Error_mc
improve_av
```

The antithetic variate reduction reduces the standard error by 75.8793%.

#### 2.Stratified Sampling
Since $1 - U_i \overset{d}{=} U_i$, we can compute
$$
h(U)=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2})}{\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2}},
$$

where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.
```{r include=FALSE}
set.seed(1)
n <- 10^5
j <- 1:n
sample_1b2 <- 4 * (sin(sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)))/sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)

Mu_SS <- mean(sample_1b2)
Diff_SS <- true_integral_1b - Mu_SS
SE_SS <- sd(sample_1b2)/sqrt(n)

rbind(Mu_SS, Diff_SS, SE_SS)
```

```{r echo=TRUE}
improve_SS <- (EstIntegral_mc - Mu_SS)/EstIntegral_mc
improve_SS
```

#### Stratified Sampling + Antithetic Variate reduction
To combine the stratified sampling with antithetic variate reduction, we calculte the mean of the following to estimations:
$$
h(U)=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2})}{\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2}},
$$
and
$$
h(1-U)=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(\frac{i-U_i}{n})^{2} + (\frac{j-U_j}{n})^2})}{\sqrt{(\frac{i-U_i}{n})^{2} + (\frac{j-U_j}{n})^2}}.
$$

```{r include=FALSE}
set.seed(1)
n <- 10^5
j <- 1:n
sample_avss_1 <- 4 * (sin(sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)))/sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)
sample_avss_2 <- 4 * (sin(sqrt(((j-x)/n)^2 + ((j-y)/n)^2)))/sqrt(((j-x)/n)^2 + ((j-y)/n)^2)
sample_avss <- (sample_avss_1 + sample_avss_2)/2

Mu_avss <- mean(sample_avss)
Diff_avss <- true_integral_1b - Mu_avss
SE_avss <- sd(sample_avss)/sqrt(n)

rbind(Mu_avss, Diff_avss, SE_avss)
```

```{r}
improve_SS <- (EstIntegral_mc - Mu_avss)/EstIntegral_mc
improve_SS
```

The following table shows the comparison between different methods:
```{r echo=FALSE}
df_1b<-data.frame(
  Method=c("Ture Value_1b","MC_1b","AV_1b","SS_1b","AVSS_1b"),
  Mean=c(true_integral_1b,EstIntegral_mc,EstIntegral_av,Mu_SS,Mu_avss),
  Diff=c("NA",Diff_mc,Mean_diff_av,Diff_SS,Diff_avss),
  Standar_error=c("NA",Std_Error_mc,Std_Error_av,SE_SS,SE_avss)
)
df_1b
```
Therefore, the antithetic variate reduction can reduce the variance by 75.8793%, and the difference between estimated mean and the true mean is the smallest. Therefore, the antithetic variate reduction is the best method to improve the accuracy.



## Question 3

### (a)

A Bessel process $R_t$ of order $\delta \geq 1$ has the SDE:

$$
dR_t = \frac{\delta - 1}{2R_t} \, dt + dW_t, \quad R_0 = 1
$$

Consider the SDE for the squared Bessel process $Y_t = R_t^2$, by Itô's formula:

$$
dY_t = 2R_t \, dR_t + \frac{1}{2} \cdot 2 \, d[R]_t, \quad \text{where } d[R]_t = dt
$$

$$
= 2R_t \left( \frac{\delta - 1}{2R_t} \, dt + dW_t \right) + dt
$$

$$
= \delta \, dt + 2 \sqrt{Y_t} \, dW_t,  \quad dW_t \sim \sqrt{dt} \cdot \mathcal{N}(0, 1)
$$

$$
Y_t = Y_0 + \int_0^t \delta \, ds + \int_0^t 2\sqrt{Y_s} \, dW_s, \quad Y_0 = R_0^2 = 1
$$

Check:

$$
\mathbb{E}(Y_t) = Y_0 + \delta t, \quad \text{since } \mathbb{E} \left( \int_0^t 2\sqrt{Y_s} \, dW_s \right) = 0
$$


```{r}
set.seed(1)
T = 10
N = 1000
dt = T/N
Rt_sim <- function(delta, N, dt){
  Y <- numeric(N+1)
  Y[1] <- 1
  for (i in 1:N){
    dW <- sqrt(dt)*rnorm(1)
    Y[i+1] <- max(Y[i] + delta*dt+ 2*sqrt(Y[i])*dW, 0)
  } 
  return(sqrt(Y))
}
t <- seq(0, T, by = dt)
delta <- c(1,2,3,4)
Rt1 <- Rt_sim(delta[1], N=1000, dt=10/1000)
Rt2 <- Rt_sim(delta[2], N=1000, dt=10/1000)
Rt3 <- Rt_sim(delta[3], N=1000, dt=10/1000)
Rt4 <- Rt_sim(delta[4], N=1000, dt=10/1000)
ymin <- min(c(Rt1, Rt2, Rt3, Rt4))
ymax <- max(c(Rt1, Rt2, Rt3, Rt4))
plot(t, Rt1, "l", col = "yellow", ylim = c(ymin, ymax), main = "Sample Paths of Bessel Process", xlab = "Time t", ylab = expression(R[t]))
lines(t, Rt2, "l", col="red")
lines(t, Rt3, "l", col="blue")
lines(t, Rt4, "l", col="green")
legend("topleft",
       legend = c(expression(delta *' = 1'), expression(delta *' = 2'), expression(delta *' = 3'), expression(delta *' = 4')),
       col = c("yellow", "red", "blue", "green"),
       lty = 1)  
```

### (b)

Generate more paths over $t \in [0, 10]$ with $\delta=1$:

```{r}
set.seed(1)
M <- 100  # Number of sample paths
t <- seq(0, T, by = dt)
# choose delta = 1, generate more paths
Rt_matrix <- sapply(1:M, function(m) Rt_sim(delta[1], N, dt)) # row: dt, col: Rt
matplot(t, Rt_matrix, type = "l", lty = 1, col = rainbow(M),
        main = "Sample Paths of Bessel Process (δ = 1)",
        xlab = "Time t", ylab = expression(R[t]))
```

Simulate the maximum for each path and calculate the mean of this maximum:

```{r}
max_Rt <- apply(Rt_matrix, 2, max) # calculate maximum for each column of Rt_matrix
hist(max_Rt, breaks = 20, prob = T, 
     main = expression("Histogram of Max " * R[t] * " over [0,10] (δ = 1)"), 
     xlab = expression("Max " * R[t]))
mean(max_Rt)
legend("topright", legend = "mean = 3.812538")
```

### (c)

Consider a Bermudan call option with the following characteristics:

- Underlying stock price: $R_t$, a Bessel process with order $\delta = 1.2$
- Strike price: $K = 2$
- Maturity: $T = 10$
- Exercisable at discrete times: $t_k = k$, for $k = 1, 2, \dots, 10$
- Exercise time: $\tau$ is a stopping time chosen from $\{t_1, t_2, \dots, t_{10}\}$
- Payoff at exercise: $(R_\tau - K)^+ = \max(R_\tau - K, 0)$

First, generate sample paths for stock price $R_t$ with $\delta=1.2$.

```{r}
set.seed(1)
delta = 1.2
K = 2
T = 10
M <- 100  # Number of sample paths
t <- seq(0, T, by = dt)
Rt_matrix <- sapply(1:M, function(m) Rt_sim(delta, N, dt)) # rows: t, cols: the mth path of Rt
matplot(t, Rt_matrix, type = "l", lty = 1, col = rainbow(M),
        main = "Sample Paths of Bessel Process (δ = 1.2)",
        xlab = "Time t", ylab = expression(R[t]))
```

Second, construct a payoff matrix for exercisable dates.
The option will be exercised when the highest payoff is reached. Assign these values to the exercise time $\tau$.

```{r}
tk_index <- (1:10)/(10/1000) + 1
Rtk <- Rt_matrix[tk_index, ]  # rows: tk, cols: the mth path of Rt
tk_payoff <- pmax(Rtk - K, 0)  # (Rt - K)+
exercise_time <- apply(tk_payoff, 2, function(payoffs) {
  idx <- which.max(payoffs)  # highest positive payoff
  if (is.na(idx)) return(NA)   # never exercised
  return(idx)  # tk index 1:10
})
# Convert index to actual time τ
tau <- ifelse(is.na(exercise_time), NA, exercise_time)  # τ in 1:10
hist(tau, prob = T, breaks = 20)
lines(density(tau[!is.na(tau)]))
summary(tau) # median = 8, mean = 6.89
```

In conclusion, the result(median=8,mean=6.89) shows that there are no incentives to exercise early. Since there are no discounting factors, most options are exercised close to maturity.

### (d)

Assuming no discounting factor, i.e. $r=0$ thus $e^{-rt}=1$, the pricing formula of the Bermuda call option is $E(max((R_{t_{i}}-K)^+, E(V_{i+1}|Fi)))$, where $V_{i+1}$ is the future value. Hence this is equivalent to find the maximum value of the payoffs all through the exercisable times, i.e. $max((R_{t_1}-K)^+,..., (R_{t_{10}}-K)^+)$

```{r}
Bermuda_payoffs <- sapply(1:M, function(m) {
  payoffs <- tk_payoff[, m]
  holding_values <- numeric(10)
  holding_values[10] <- payoffs[10]
  for (t in 9:1) {
    if (payoffs[t] > 0) {
      immediate_exercise <- payoffs[t]
      future_payoff <- holding_values[t + 1]
      holding_values[t] <- max(immediate_exercise, future_payoff)
    } else {
      holding_values[t] <- holding_values[t + 1]
    }
  }
  return(holding_values[1])
})
Bermuda_price <- mean(Bermuda_payoffs)
cat("Bermuda call price =", Bermuda_price, "\n")
```

The European call option price $= E((R_{10} - K)^+)$.

```{r}
# European call option price = E((R10-K)+)
European_payoff <- tk_payoff[10,]
European_price <- mean(European_payoff)
cat("European call price =", European_price, "\n")
```

Calculate how much more expensive the Bermuda call option is than a European
call option:

```{r}
diff_price = Bermuda_price - European_price
cat("Extra cost of Bermuda over European =", diff_price, "\n")
```

### (e)

#### (1) No proper modeling of expected future values:
We used a single “average of future payoffs” or just “the next time step’s payoff,” which might be a coarse approximation.
The more accurate method in continuous models is to do a regression across all paths’ next‐step (or multiple‐step) payoffs on the underlying Rt.

#### (2) Assumption of the discouting factor is not realistic:
In real-world financial markets, option pricing typically incorporates a positive risk-free interest rate $r > 0$, leading to exponential discounts of future payoffs. 

When discounting is present, the option value at an exercise time $t_i$ is given by:

$$
V_i = \max\left( (R_{t_i} - K)^+, \ e^{-r \Delta t} \cdot \mathbb{E}[V_{i+1} \mid \mathcal{F}_{t_i}] \right)
$$

The second term represents the discounted expected continuation value, where $\Delta t$ is the time between exercise opportunities, and $ \mathcal{F}_{t_i} $ denotes the available information at time $t_i$.

In contrast to a zero-interest-rate setting where the continuation value is not discounted, a positive $r$ reduces the present value of waiting. This provides a stronger incentive for early exercise and typically increases the premium of the Bermuda option over its European counterpart.

#### (3) Limited number of paths:
If the number of paths is relatively small (e.g. we use M=100 in this case), the variance of the estimator might be high.
Checking for confidence intervals or repeating with a bigger M can highlight whether results have converged.

#### (4) Issues arising from the direct simulation of Rt:

$$
    R_t \to 0 \quad \Rightarrow \quad Y_t = R_t^2 \to 0
$$ 
To avoid taking the square root of negative values when simulating $R_t = \sqrt{Y_t}$, we applied a non-negativity constraint in the simulation of $Y_t$:

$$
Y_{i+1} = \max\left( Y_i + \delta \cdot \Delta t + 2 \sqrt{Y_i} \cdot \Delta W_i,\ 0 \right)
$$

This ensures that $Y_t \geq 0$, so that $R_t = \sqrt{Y_t}$ is always real-valued.

#### (5) Variance Reduction:

Control Variate Method:

The price of the Bermuda call option is equal to $max((R_{t_1}-K)^+,..., (R_{t_{10}}-K)^+)$, which is equivalent to $max(max(R_t)-K,\ 0)$. 

Let $X = max(max(R_t)-K,\ 0), \ for\ t=1,2,...,10$. Set $\tilde{X} = max(R_t)-K$, where we can use similar method in (b) to calculate $max(R_t)$, then $Y = \tilde{X} - E(\tilde{X})$. Check: $E(Y) = 0$.

```{r}
max_Rtk <- apply(Rtk, 2, max)
X_tilde = max_Rtk - K
X = pmax(X_tilde, 0)
Y = X_tilde - mean(X_tilde)
mean(Y) # check E(Y)=0
```

Calculate $c=-\frac{Cov(X,Y)}{Var(Y)}$:

```{r}
c <- -cov(X,Y)/var(Y)
c
```

Then $\hat{\theta}=X + cY$, calculate $Var(\hat{\theta})$ and compare it to $Var(X)$:
```{r}
theta_hat = X + c*Y
cat(var(X), var(theta_hat), "\n",  
    "Improved by",round((var(X)-var(theta_hat))/var(X) * 100, digits = 2), "%.")
```

Other attempts for variance reduction:

1. Antithetic Variates Method: unavailable since $R_t$ is not a monotone function.
2. Stratify Sampling: unavailable since an inverse transformed estimator is needed.
3. Conditioning Sampling: unavailable since $R_t$ is the only variable.
4. Importance Sampling: unavailable since the distribution of $R_t$ is not given.



## Question 4

### (a)

First, we create a function that models one game and builds a deck with numbers 1-8 (with four numbers repeated).
Then, we randomly deal 4 cards to each player.
The player who has the lowest sum will be the winner of the game.

```{r}
sim_game <- function() {
  nums <- 1:8
  repeat_nums <- sample(nums, size=4, replace = FALSE)
  my_deck <- c(nums, repeat_nums)
  
  # Draw cards without replacement
  cards_drawn <- sample(my_deck, size=8, replace = FALSE)
  
  # Split cards between players
  player1_cards <- cards_drawn[1:4]
  player2_cards <- cards_drawn[5:8]
  
  # Find totals
  p1_total <- sum(player1_cards)
  p2_total <- sum(player2_cards)
  
  # Figure out who wins and calculate payoff
  result <- sign(p2_total - p1_total)
  
  if (result > 0) {  # Player 1 wins
    pay <- min(player1_cards)
    return(c(pay, -pay))
  } else if (result < 0) {  # Player 2 wins
    pay <- min(player2_cards)
    return(c(-pay, pay))
  } else {  # Nobody wins 
    return(c(0, 0))
  }
}

# Run simulation 100,000 times
set.seed(42)
sims <- 100000
all_results <- matrix(0, nrow = sims, ncol = 2)
for (i in 1:sims) all_results[i,] <- sim_game()

# Find unique outcomes for histograms
p1_outcomes <- sort(unique(all_results[,1]))
p2_outcomes <- sort(unique(all_results[,2]))

# Make histograms
par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
hist(all_results[,1], breaks = p1_outcomes, prob = TRUE, 
     main = "Player 1 Payoffs", xlab = "Amount")
hist(all_results[,2], breaks = p2_outcomes, prob = TRUE, 
     main = "Player 2 Payoffs", xlab = "Amount")

# Calculate expected values
avg_p1 <- mean(all_results[,1])
avg_p2 <- mean(all_results[,2])
cat("\nPlayer 1 expected value:", round(avg_p1, 4))
cat("\nPlayer 2 expected value:", round(avg_p2, 4))
```

Conclusion: 
The payoff distributions for both players are mirror images of each other.\
The most common outcomes are -1, 0, 1, and 2, with less frequent outcomes at -3, -2, 3, and 4.\ 
The expected values (Player 1: -0.0077, Player 2: 0.0077) are essentially zero.\ 
The game appears to be fair, with neither player having a significant advantage.\ 
Ties (payoff of 0) occur frequently, as shown by the peak at 0 in both histograms. 

### (b) 
```{r}
# Create new function where Player 1 always gets their chosen card
sim_game_modified <- function(chosen) {
  nums <- 1:8
  repeat_nums <- sample(nums, size=4, replace = FALSE)
  my_deck <- c(nums, repeat_nums)
  
  # Player 1 gets chosen card plus 3 random cards
  # Player 2 gets 4 random cards
  if (chosen %in% my_deck) {
    idx <- which(my_deck == chosen)[1]
    my_deck <- my_deck[-idx]
  } else {
    return(sim_game_modified(chosen))
  }
  
  cards_drawn <- sample(my_deck, size=7, replace = FALSE)
  player1_cards <- c(chosen, cards_drawn[1:3])
  player2_cards <- cards_drawn[4:7]
  
  p1_total <- sum(player1_cards)
  p2_total <- sum(player2_cards)
  
  if (p1_total < p2_total) {
    pay <- min(player1_cards)
    return(c(pay, -pay))
  } else if (p2_total < p1_total) {
    pay <- min(player2_cards)
    return(c(-pay, pay))
  } else {
    return(c(0, 0))
  }
}

# Test each possible card choice (1-8) with 50,000 simulations
set.seed(42)
sims <- 50000
card_choices <- 1:8
choice_results <- matrix(0, nrow = 8, ncol = 2)
colnames(choice_results) <- c("Exp Value", "Win %")

for (i in 1:8) {
  test_results <- matrix(0, nrow = sims, ncol = 2)
  for (j in 1:sims) test_results[j,] <- sim_game_modified(i)
  
  choice_results[i,] <- c(
    mean(test_results[,1]),
    mean(test_results[,1] > 0) * 100
  )
}

# Print results
cat("\nCard choice results (Expected Value, Win %):")
for (i in 1:8) {
  cat("\nCard", i, ":", round(choice_results[i,1], 4), 
      ", ", round(choice_results[i,2], 1), "%")
}

best <- which.max(choice_results[,1])
cat("\n\nBest choice for Player 1: Card", best)
cat("\nThis gives expected value of", round(choice_results[best,1], 4), "per game")
cat("\nThe fair game had expected value around 0, so this is definitely better")
```

Conclusion: \
The player A should choose card 2.
It gives the highest expected value of payoff equal to 0.518 per game. \
While choosing card 1 offers the highest win probability (70.9%), it only yields an expected value of payoff equal to 0.294. \
Card 2 provides a better balance of win probability (63.4%) and higher payoff when winning. \
As card numbers increase, both expected values of payoff and win probabilities decrease steadily. \

Player A now has a significant advantage.