---
title: "Q1-Q2"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
```{r}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
plot(grid_t, exact_values_C, type = 'l')
```

### Monte Carlo Methods
```{r}
n <- 10^6
```

#### Transformation 1
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```


##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample)
  
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_4 <- numeric(length(grid_t))
Diff_4 <- numeric(length(grid_t))
Std_Error_4 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_4[t] <- mean(sample)
  
  Diff_4[t] <- exact_formula_C(t) - EstIntegral_4[t]
  Std_Error_4[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_4 = mean(Diff_4), mean_SE_4 = mean(Std_Error_4))
```

#### Transformation 3
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```
##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_4 <- numeric(length(grid_t))
Diff_4 <- numeric(length(grid_t))
Std_Error_4 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_4[t] <- mean(sample)
  
  Diff_4[t] <- exact_formula_C(t) - EstIntegral_4[t]
  Std_Error_4[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_4 = mean(Diff_4), mean_SE_4 = mean(Std_Error_4))
```

According to the transformation methods above, transformation 3 has the smallest standard error. Thus we do the variance reduction based on transformation 3.
#### 1.Antithetic Variate reduction
U=1-U
```{r}
t<-1
u <- runif(n)
samplex <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
sampley <- ((sin(pi/2*(1-u)))^t + (cos(pi/2*(1-u)))^t) * pi/2

mu_antithetic<-mean((samplex+sampley)/2)
Diff <- exact_formula_C(t)-mu_antithetic

SE_antithetic<-sd(0.5*(samplex+sampley)/sqrt(n))
cbind(mu_antithetic, Diff, SE_antithetic)
```


## b)
### Calculate the true value of the double integral
```{r}
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
true_integral <- result$value
true_integral
```

#### Alternatively, using another package:
```{r}
# install.packages('pracma')
library(pracma)

f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r) 
}

dblquad(f, -1, 1, -1, 1)
```

### Pure Monte Carlo Methods
```{r}
n <- 10^5
x <- runif(n)
y <- runif(n)
sample <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)

EstIntegral <- mean(sample)
Diff <- true_integral - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```

# Question 2
## a)
### Generate various values of m, nu, lambda and alpha
```{r}
m_list <- c(2, 10)
nu_list <- c(-1, 1)
lambda_list <- c(-1, 1)
alpha_list <- c(1, 10)

groups <- expand.grid(m_list, nu_list, lambda_list, alpha_list)

m <- groups$Var1
nu <- groups$Var2
lambda <- groups$Var3
alpha <- groups$Var4

# convert rows to list of vectors
group_list <- split(as.matrix(groups), LETTERS[1:nrow(groups)])
group_list
```


```{r}
# m <- 2
# nu <- -1
# lambda <- -1
# alpha <- 1
```



A temporary check function:
```{r}
theoretical_mean <- lambda - alpha*nu/(2*(m - 1))
theoretical_var <- alpha^2/(2*m - 3) * (1 + nu^2/(4*(m - 1)^2))
```


```{r}
Q2_summary <- data.frame(m, nu, lambda, alpha, theoretical_mean, theoretical_var)
```



### With transformation
$$
u = tan^{-1}(\frac{x - \lambda}{\alpha}), u \in [-\frac{\pi}{2}, \frac{\pi}{2}]
$$
We get
$$
f(u) \propto (cos(u))^{2m}e^{-\nu u}
$$




#### Envelope function 1 (uniform distribution)
$$
g(y) = \frac{1}{\pi}, Y \sim Unif[-\frac{\pi}{2}, \frac{\pi}{2}]
$$

```{r}
# Generate Y
n <- 10^5
Y <- runif(n) * pi - pi/2
```

```{r}
# Generate the sample using AR
U <- runif(n)

sample_envelope_1 <- lapply(1:16, function(i) {
  m <- m[i]
  nu <- nu[i]
  
  temp <- atan(-nu / (2 * m))
  accepted_Y <- Y[U < 
    (cos(Y))^(2 * m) * exp(-nu * Y) / (cos(temp)^(2 * m) * exp(-nu * temp))]
})

plot(density(sample_envelope_1[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_1 <- lapply(1:16, function(i){tan(sample_envelope_1[[i]]) * alpha[i] + lambda[i]})
# X_envelope_1[[1]][1:10]
# sample_envelope_1[[1]][1:10]

envelop_1_empirical_mean <- unlist(lapply(1:16, function(i) mean(X_envelope_1[[i]])))
envelop_1_empirical_var <- unlist(lapply(1:16, function(i) var(X_envelope_1[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_1_empirical_acceptance <- unlist(lapply(1:16, function(i) length(sample_envelope_1[[i]])/10^5))
```

```{r}
# add the results to the summary
library(tidyverse)
Q2_summary <- Q2_summary %>% 
  mutate(envelop_1_empirical_mean, envelop_1_empirical_var, envelope_1_empirical_acceptance)

Q2_summary
```

#### Envelope function 2 (exponential)
$$
g(y) = \frac{\nu e^{-\nu y}}{e^{\frac{\pi}{2}\nu} - e^{-\frac{\pi}{2}\nu}}
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = -\frac{1}{2}log[U \times (e^{\frac{-\pi}{2}\nu} - e^{\frac{\pi}{2}\nu}) + e^{\frac{\pi}{2}\nu}]
$$

```{r}
# Generate Y
n <- 10^5
U <- runif(n)

Y <- lapply(1:16, function(i){-0.5*log(U*(exp(-pi/2 * nu[i]) - exp(pi/2 * nu[i])) + exp(pi/2 * nu[i]))})
```

```{r}
# Generate the sample using AR
U <- runif(n)

sample_envelope_2 <- lapply(1:16, function(i){
  Y[[i]][U < (cos(Y[[i]]))^{2*m[i]}]
})

plot(density(sample_envelope_2[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_2 <- lapply(1:16, function(i) tan(sample_envelope_2[[i]]) * alpha[i] + lambda[i])

envelop_2_empirical_mean <- unlist(lapply(1:16, function(i) mean(X_envelope_2[[i]])))
envelop_2_empirical_var <- unlist(lapply(1:16, function(i) var(X_envelope_2[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_2_empirical_acceptance <- unlist(lapply(1:16, function(i) length(sample_envelope_2[[i]])/10^5))
```
```{r}
# update Q2_summary
Q2_summary <- Q2_summary %>% 
  mutate(envelop_2_empirical_mean, envelop_2_empirical_var, envelope_2_empirical_acceptance)

Q2_summary
```


#### Envelope function 3 (cosine)
$$
g(y) = \frac{1}{2}cos(y)
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = sin^{-1}(2U - 1)
$$

```{r}
# Generate Y
n <- 10^5
U <- runif(n)

Y <- asin(2*U-1)
```

```{r}
# Generate the sample using AR
U <- runif(n)

temp <- atan(-nu/(2*m-1))

sample_envelope_3 <- lapply(1:16, function(i){
  temp <- temp[i]
  m <- m[i]
  nu <- nu[i]
  
  Y[U < (cos(Y)^{2*m-1}*exp(-nu*Y))/(cos(temp)^{2*m-1}*exp(-nu*temp))]
})

plot(density(sample_envelope_3[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_3 <- lapply(1:16, function(i) tan(sample_envelope_3[[i]]) * alpha[i] + lambda[i])

envelop_3_empirical_mean <- unlist(lapply(1:16, function(i) mean(X_envelope_3[[i]])))
envelop_3_empirical_var <- unlist(lapply(1:16, function(i) var(X_envelope_3[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_3_empirical_acceptance <- unlist(lapply(1:16, function(i) length(sample_envelope_3[[i]])/10^5))
```

```{r}
# update Q2_summary
Q2_summary <- Q2_summary %>% 
  mutate(envelop_3_empirical_mean, envelop_3_empirical_var, envelope_3_empirical_acceptance)
```

### Without transformation
#### Envelope function 4 (Cauchy)

```{r}
# Generate Y
n <- 10^5
Y <- lapply(1:16, function(i) alpha[i] * tan(pi*runif(n) - pi/2) + lambda[i])
```

```{r}
# Generate the sample using AR
U <- runif(n)

temp_1 <- lapply(1:16, function(i) (Y[[i]] - lambda[i])/alpha[i])
temp_2 <- nu/(2*(1 - m))

sample_envelope_4 <- lapply(1:16, function(i){
  Y <- Y[[i]]
  temp_1 <- temp_1[[i]]
  temp_2 <- temp_2[i]
  nu <- nu[i]
  m <- m[i]
  
  Y[U < ((1 + temp_1^2)^{-m} * exp(-nu * atan(temp_1)))/((1 + temp_2^2)^{1-m} * exp(-nu*atan(temp_2)))]
})

summary(sample_envelope_4)

plot(density(sample_envelope_4[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_4 <- sample_envelope_4

envelop_4_empirical_mean <- unlist(lapply(1:16, function(i) mean(X_envelope_4[[i]])))
envelop_4_empirical_var <- unlist(lapply(1:16, function(i) var(X_envelope_4[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_4_empirical_acceptance <- unlist(lapply(1:16, function(i) length(sample_envelope_4[[i]])/10^5))
```

```{r}
# update summary_Q2
Q2_summary <- Q2_summary %>% 
  mutate(envelop_4_empirical_mean, envelop_4_empirical_var, envelope_4_empirical_acceptance)

colnames(Q2_summary)
```


#### Envelope function 5 (t distribution)
```{r}
# Generate Y
n <- 10^5
k <- 2*m - 1

Y <- lapply(1:16, function(i){
  rt(n, df = k[i]) * alpha[i]/sqrt(k[i]) + lambda[i]
})
```

```{r}
# Generate the sample using AR
U <- runif(n)

sample_envelope_5 <- lapply(1:16, function(i){
  Y <- Y[[i]]
  nu <- nu[i]
  lambda <- lambda[i]
  alpha <- alpha[i]
  
  Y[U < exp(-nu*atan((Y - lambda)/alpha) - nu*pi/2)]
})

plot(density(sample_envelope_5[[1]]), type = 'l')
```

```{r}
# Check mean and variance
X_envelope_5 <- sample_envelope_5

envelop_5_empirical_mean <- unlist(lapply(1:16, function(i) mean(X_envelope_5[[i]])))
envelop_5_empirical_var <- unlist(lapply(1:16, function(i) var(X_envelope_5[[i]])))
```


```{r}
# empirical rate of acceptance
envelope_5_empirical_acceptance <- unlist(lapply(1:16, function(i) length(sample_envelope_5[[i]])/10^5))
```

```{r}
# update summary_Q2
Q2_summary <- Q2_summary %>% 
  mutate(envelop_5_empirical_mean, envelop_5_empirical_var, envelope_5_empirical_acceptance)
```


### Export summary_Q2
```{r}
write.csv(Q2_summary, 'Q2_summary.csv', row.names = F)
```



