---
title: "Q1-Q2"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
```{r}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
plot(grid_t, exact_values_C, type = 'l')
```

### Monte Carlo Methods
```{r}
n <- 10^6
```

#### Transformation 1
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```


##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample)
  
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_2 <- numeric(length(grid_t))
Diff_2 <- numeric(length(grid_t))
Std_Error_2 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_2[t] <- mean(sample)
  
  Diff_2[t] <- exact_formula_C(t) - EstIntegral_2[t]
  Std_Error_2[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_2 = mean(Diff_2), mean_SE_2 = mean(Std_Error_2))
```

#### Transformation 3
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```
##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_3 <- numeric(length(grid_t))
Diff_3 <- numeric(length(grid_t))
Std_Error_3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_3[t] <- mean(sample)
  
  Diff_3[t] <- exact_formula_C(t) - EstIntegral_3[t]
  Std_Error_3[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_3 = mean(Diff_3), mean_SE_3 = mean(Std_Error_3))
```

## b)
### Calculate the true value of the double integral
```{r}
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
true_integral <- result$value
```

#### Alternatively, using another package:
```{r}
# install.packages('pracma')
library(pracma)

f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r) 
}

dblquad(f, -1, 1, -1, 1)
```

### Pure Monte Carlo Methods
```{r}
n <- 10^5
x <- runif(n)
y <- runif(n)
sample <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)

EstIntegral <- mean(sample)
Diff <- true_integral - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```

# Question 2
## a)
### Generate various values of m, nu, lambda and alpha
```{r}
# m <- c(2, 10)
# nu <- c(-1, 1)
# lambda <- c(-1, 1)
# alpha <- c(1, 10)
# 
# groups <- expand.grid(m, nu, lambda, alpha)
# 
# # convert rows to list of vectors
# group_list <- split(as.matrix(groups), 1:nrow(groups))
# #group_list
```


```{r}
m <- 2
nu <- -1
lambda <- -1
alpha <- 1
```

A temporary check function:
```{r}
check_mean_var <- function(sample){
  mean <- lambda - alpha*nu/(2*(m - 1))
  var <- alpha^2/(2*m - 3) * (1 + nu^2/(4*(m - 1)^2))
  cat(mean, var)
}
```


With transformation
$$
u = tan^{-1}(\frac{x - \lambda}{\alpha}), u \in [-\frac{\pi}{2}, \frac{\pi}{2}]
$$
We get
$$
f(u) \propto (cos(u))^{2m}e^{-\nu u}
$$




#### Envelop function 1 (uniform distribution)
$$
g(y) = \frac{1}{\pi}, Y \sim Unif[-\frac{\pi}{2}, \frac{\pi}{2}]
$$

```{r}
# Generate Y
n <- 10^5
Y <- runif(n) * pi - pi/2
```

```{r}
c_over_K_1 <- pi * (cos(atan(-nu/(2*m))))^{2*m} * exp(-nu*atan(-nu/(2*m)))
c_over_K_1
```

```{r}
# Generate the sample using AR
U <- runif(n)

temp <- atan(-nu/(2*m))

sample_envelope_1 <- Y[U < 
                      ((cos(Y))^{2 * m} * exp(-nu * Y))/(cos(temp)^{2*m} * exp(-nu * temp))]

plot(density(sample_envelope_1), type = 'l')
```


```{r}
# Produce X, and check mean and variance
X_envelope_1 <- tan(sample_envelope_1) * alpha + lambda
print(c(mean(X_envelope_1), var(X_envelope_1)))
check_mean_var(X_envelope_1)
```
```{r}
# empirical rate of acceptance
length(sample_envelop_1)/n
```

#### Envelope function 2 (cosine squared)
```{r}

```


#### Envelope function 3 (exponential)
$$
g(y) = \frac{\nu e^{-\nu y}}{e^{\frac{\pi}{2}\nu} - e^{-\frac{\pi}{2}\nu}}
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = -\frac{1}{2}log[U \times (e^{\frac{-\pi}{2}\nu} - e^{\frac{\pi}{2}\nu}) + e^{\frac{\pi}{2}\nu}]
$$

```{r}
# Generate Y
n <- 10^5
U <- runif(n)

Y <- -0.5*log(U*(exp(-pi/2 * nu) - exp(pi/2 * nu)) + exp(pi/2 * nu))
```

```{r}
# Generate the sample using AR
U <- runif(n)

sample_envelope_2 <- Y[U < (cos(Y))^{2*m}]

plot(density(sample_envelope_2), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_2 <- tan(sample_envelope_2) * alpha + lambda
print(c(mean(X_envelope_2), var(X_envelope_2)))
check_mean_var(X_envelope_2)
```

#### Envelope function 2 (cosine squared)


