---
title: "Q1-Q2"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
```{r}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
plot(grid_t, exact_values_C, type = 'l')
```

### Monte Carlo Methods
```{r}
n <- 10^6
```

#### Transformation 1
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```


##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample)
  
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_2 <- numeric(length(grid_t))
Diff_2 <- numeric(length(grid_t))
Std_Error_2 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_2[t] <- mean(sample)
  
  Diff_2[t] <- exact_formula_C(t) - EstIntegral_2[t]
  Std_Error_2[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_2 = mean(Diff_2), mean_SE_2 = mean(Std_Error_2))
```

#### Transformation 3
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```
##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_3 <- numeric(length(grid_t))
Diff_3 <- numeric(length(grid_t))
Std_Error_3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_3[t] <- mean(sample)
  
  Diff_3[t] <- exact_formula_C(t) - EstIntegral_3[t]
  Std_Error_3[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_3 = mean(Diff_3), mean_SE_3 = mean(Std_Error_3))
```


## b)
```{r}
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
result$value


```

##### Alternatively, using another package:
```{r}
# install.packages('pracma')
library(pracma)

f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r) 
}

dblquad(f, -1, 1, -1, 1)
```

