---
title: "Q1"
output: html_document
date: "2025-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
```{r}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
plot(grid_t, exact_values_C, type = 'l')
```

### Monte Carlo Methods
```{r}
n <- 10^6
```

#### Transformation 1
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```


##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample)
  
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_4 <- numeric(length(grid_t))
Diff_4 <- numeric(length(grid_t))
Std_Error_4 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_4[t] <- mean(sample)
  
  Diff_4[t] <- exact_formula_C(t) - EstIntegral_4[t]
  Std_Error_4[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_4 = mean(Diff_4), mean_SE_4 = mean(Std_Error_4))
```

#### Transformation 3
##### Investigate the performance of integral for a fixed value of t
```{r}
t <- 1
u <- runif(n)
sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
EstIntegral <- mean(sample)
  
Diff <- exact_formula_C(t) - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```
##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_4 <- numeric(length(grid_t))
Diff_4 <- numeric(length(grid_t))
Std_Error_4 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_4[t] <- mean(sample)
  
  Diff_4[t] <- exact_formula_C(t) - EstIntegral_4[t]
  Std_Error_4[t] <- sd(sample)/sqrt(n)
}

rbind(mean_diff_4 = mean(Diff_4), mean_SE_4 = mean(Std_Error_4))
```

According to the transformation methods above, transformation 3 has the smallest standard error. Thus we do the variance reduction based on transformation 3.

The antithetic variate reduction is not applicable since the function is not monotone.

#### 1.Control variates method
```{r}
t <- 1
u <- runif(n)
samplex <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
sampley <- ((sin(pi/2*(1-u)))^t + (cos(pi/2*(1-u)))^t) * pi/2

mu_antithetic<-mean((samplex+sampley)/2)
Diff <- exact_formula_C(t)-mu_antithetic

SE_antithetic<-sd(0.5*(samplex+sampley))/sqrt(2*n)
cbind(mu_antithetic, Diff, SE_antithetic)
```

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
EstIntegral_4 <- numeric(length(grid_t))
Diff_4 <- numeric(length(grid_t))
Std_Error_4 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  samplex <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  sampley <- ((sin(pi/2*(1-u)))^t + (cos(pi/2*(1-u)))^t) * pi/2
  
  mu_antithetic[t] <- mean((samplex+sampley)/2)
  
  Diff_4[t] <- exact_formula_C(t) - mu_antithetic[t]
  Std_Error_4[t] <- sd(0.5*(samplex+sampley))/sqrt(2*n)
}

rbind(mean_diff_4 = mean(mu_antithetic), mean_SE_4 = mean(Std_Error_4))
```

## b)
### Calculate the true value of the double integral
```{r}
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
true_integral <- result$value
true_integral
```

#### Alternatively, using another package:
```{r}
# install.packages('pracma')
library(pracma)

f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r) 
}

dblquad(f, -1, 1, -1, 1)
```

### Pure Monte Carlo Methods
```{r}
n <- 10^5
x <- runif(n)
y <- runif(n)
sample <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)

EstIntegral <- mean(sample)
Diff <- true_integral - EstIntegral
Std_Error <- sd(sample)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```
#### 1.Antithetic Variate reduction
```{r}
n <- 10^5
x <- runif(n)
y <- runif(n)
sample1 <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)
sample2 <- 4 * sin(sqrt((1-x)^2 + (1-y)^2))/sqrt((1-x)^2 + (1-y)^2)
EstIntegral <- mean((sample1+sample2)/2)
Diff <- true_integral - EstIntegral
Std_Error <- sd((sample1+sample2)/2)/sqrt(n)

rbind(EstIntegral, Diff, Std_Error)
```