---
title: "Q1"
output: html_document
date: "2025-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
```{r}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
plot(grid_t, exact_values_C, type = 'l')
```

### Monte Carlo Methods
```{r}
n <- 10^6
```

#### Transformation 1

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
set.seed(1)
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_1 <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample_1)
  
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample_1)/sqrt(n)
}

rbind(mean_1=mean(EstIntegral_1),mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
set.seed(1)
EstIntegral_2 <- numeric(length(grid_t))
Diff_2 <- numeric(length(grid_t))
Std_Error_2 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_2 <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_2[t] <- mean(sample_2)
  
  Diff_2[t] <- exact_formula_C(t) - EstIntegral_2[t]
  Std_Error_2[t] <- sd(sample_2)/sqrt(n)
}

rbind(mean_2 = mean(EstIntegral_2), mean_diff_2 = mean(Diff_2), mean_SE_2 = mean(Std_Error_2))
```

#### Transformation 3

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
set.seed(1)
EstIntegral_3 <- numeric(length(grid_t))
Diff_3 <- numeric(length(grid_t))
Std_Error_3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_3[t] <- mean(sample)
  
  Diff_3[t] <- exact_formula_C(t) - EstIntegral_3[t]
  Std_Error_3[t] <- sd(sample)/sqrt(n)
}

rbind(mean_3=mean(EstIntegral_3), mean_diff_3 = mean(Diff_3), SE_3 = mean(Std_Error_3))

```

According to the transformation methods above, transformation 3 has the smallest standard error. Thus we do the variance reduction based on transformation 3.

The antithetic variate reduction is not applicable since the function is not monotone. Then we try control variates method first.

#### 1.Control variates method

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
set.seed(1)
EstIntegral_4 <- numeric(length(grid_t))
Diff_4 <- numeric(length(grid_t))
Std_Error_4 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  X <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  Y <- pi^3/48*t-pi^3/16*t*u^2
  
  c <- -cov(X,Y)/var(Y)
  X_cv <- X+c*Y
  mean_cv <- mean(X_cv)
  SE_cv <- sd(X_cv)/sqrt(n)
}

rbind(mean = mean_cv, mean_SE = mean(SE_cv))
improve_1 <- abs(mean(SE_cv)-mean(Std_Error_3))/mean(Std_Error_3)
cat("The control variate method improves the accuarcy by",improve_1)
```

## b)
### Calculate the true value of the double integral
```{r}
set.seed(1)
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
true_integral_1b <- result$value
true_integral_1b
```

#### Alternatively, using another package:
```{r}
set.seed(1)
# install.packages('pracma')
library(pracma)

f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r) 
}

dblquad(f, -1, 1, -1, 1)
```

### Pure Monte Carlo Methods
```{r}
set.seed(1)

n <- 10^5
x <- runif(n)
y <- runif(n)
sample_1b <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)

EstIntegral_1b <- mean(sample_1b)
Diff_1b <- true_integral_1b - EstIntegral_1b
Std_Error_1b <- sd(sample)/sqrt(n)

rbind(EstIntegral_1b, Diff_1b, Std_Error_1b)
```

#### 1.Antithetic Variate reduction
```{r}
set.seed(1)
n <- 10^5
x <- runif(n)
y <- runif(n)
sample1 <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)
sample2 <- 4 * sin(sqrt((1-x)^2 + (1-y)^2))/sqrt((1-x)^2 + (1-y)^2)
EstIntegral_1b1 <- mean((sample1+sample2)/2)
Diff_1b1 <- true_integral_1b - EstIntegral_1b1
Std_Error_1b1 <- sd((sample1+sample2)/2)/sqrt(n)

rbind(EstIntegral_1b1, Diff_1b1, Std_Error_1b1)
improve_2 <- abs(Std_Error_1b1-Std_Error_1b)/Std_Error_1b
cat("The antithetic variate reduction improves the accuarcy by",improve_2)
```

#### 2.Stratified sampling
```{r}
set.seed(1)
n <- 10^5
u <- runif(n)
j <- 1:n
sample_1b2 <- 4 * (sin(sqrt(((u+j-1)/n)^2 + ((u+j-1)/n)^2)))/sqrt(((u+j-1)/n)^2 + ((u+j-1)/n)^2)

Mu_SS <- mean(sample_1b2)
Diff <- true_integral - Mu_SS
SE_SS <- sd(sample_1b2)/sqrt(n)

rbind(Mu_SS, Diff, SE_SS)
improve_3 <- abs(SE_SS-Std_Error_1b)/Std_Error_1b
cat("The statified sampling improves the accuarcy by",improve_3)
```

Therefore, the antithetic variate reduction can improve the accuarcy by 86.47548 %, which is the best.