---
title: "Q1"
output:
  pdf_document: default
  html_document: default
date: "2025-04-07"
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
The following diagram shows the exact values of C(t) for t from 1 to 100.
```{r eval=FALSE, include=FALSE}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
plot(grid_t, exact_values_C, type = 'l')
```

### Monte Carlo Methods
```{r}
n <- 10^6
```

#### Transformation 1

For transformation 1, we use $u=sinx$, then $x=arcsin(u)$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{x=\arcsin(u)}{=} \int_{0}^{1} {u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[{u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}}\right] \overset{MC}{\approx} \sum_{i=1}^{N}[u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}]
$$

##### Using for loop to investigate the performance of integral across different values of t:
```{r}
set.seed(1)
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_1 <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample_1)
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample_1)/sqrt(n)
}

rbind(mean_1=mean(sample_1),mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2
For transformation 2, we use $u=cosx$, then $x=\arccos(u)$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{ x=\arccos(u)}{=} \int_{0}^{1} {u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[{u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}}\right] \overset{MC}{\approx} \sum_{i=1}^{N}[u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}]
$$
##### Using for loop to investigate the performance of integral across different values of t:
```{r echo=FALSE}
set.seed(1)
EstIntegral_2 <- numeric(length(grid_t))
Diff_2 <- numeric(length(grid_t))
Std_Error_2 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_2 <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_2[t] <- mean(sample_2)
  Diff_2[t] <- exact_formula_C(t) - EstIntegral_2[t]
  Std_Error_2[t] <- sd(sample_2)/sqrt(n)
}

rbind(mean_2=mean(sample_2),mean_diff_2 = mean(Diff_2), mean_SE_2 = mean(Std_Error_2))
```
According to the above results, transformation 1 and 2 have exactly the same mean and standard error.

#### Transformation 3
For transformation 3, we use $u=\frac{2}\pi*x$, then $x=\frac{\pi}2u$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{x=\frac{\pi}2u}{=} \int_{0}^{1} {[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]\frac{\pi}2}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[\frac{\pi}2[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]\right] \overset{MC}{\approx} \sum_{i=1}^{N}\frac{\pi}2[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]
$$
##### Using for loop to investigate the performance of integral across different values of t:
```{r echo=TRUE}
set.seed(1)
EstIntegral_3 <- numeric(length(grid_t))
Diff_3 <- numeric(length(grid_t))
Std_Error_3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_3 <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_3[t] <- mean(sample_3)
  Diff_3[t] <- exact_formula_C(t) - EstIntegral_3[t]
  Std_Error_3[t] <- sd(sample_3)/sqrt(n)
}

rbind(mean_diff_3 = mean(Diff_3), SE_3 = mean(Std_Error_3))
```

The following table shows the comparison between different methods:
```{r echo=TRUE}
df_ts<-data.frame(
  Method=c("Transformation 1","Transformation 2","Transformation 3"),
  Mean=c(mean(sample_1),mean(sample_2),mean(sample_3)),
  Mean_diff=c(mean(Diff_1),mean(Diff_2),mean(Diff_3)),
  Standar_error=c(mean(Std_Error_1),mean(Std_Error_2),mean(Std_Error_3))
)
df_ts
```

According to the transformation methods above, transformation 3 has the smallest difference between estimated mean and actual mean, and also the smallest standard error. Thus we do the variance reduction based on transformation 3.

The antithetic variate reduction is not applicable since the function is not monotone. Then we try control variates method first.

#### 1.Control variates method
We have to estimate
$$
  \theta = \mathbb{E}(X) = \int_{0}^{1}{\frac{\pi}2[(\sin(\frac{\pi}2u))^{t}+(\cos(\frac{\pi}2))^{t}]}\,du,\quad\\
  \text{where}\quad X=g(U)=\frac{\pi}2[(\sin(\frac{\pi}2u))^{t}+(\cos(\frac{\pi}2u)^{t}],\quad U \overset{d}{\sim} \mathcal{U}(0,1).
$$
We want to find a random variable $Y$ with mean $0$ and a good constant $c$ such that $\mathbb{E}(X+cY) \approx \theta$.

We consider the Taylor polynomial of $g$ of degree 2 at 0:
$$
    T_1(u) = g(0) + g^{'}(0)(u-0) + \frac{1}{2}g^{''}(0)(u-0)^2 + O(u^3) = \frac{\pi}2 -\frac{\pi^{3}t}{16}u^{2},\\
     \quad\text{where}\quad g^{'}(u) =  \frac{\pi^2}{4}[t(\sin(\frac{\pi}{2}u))^{t-1}\cos(\frac{\pi}{2}u))-t(\cos(\frac{\pi}{2}u))^{t-1}\sin(\frac{\pi}{2}u))],\quad g^{'}(0) =0\\
    \quad g^{''}(u) =\frac{\pi^4}{4}[t(t-1)\frac{\pi}{2}\sin(\frac{\pi}{2}u))^{t-2}\cos^{2}(\frac{\pi}{2})-\frac{\pi}{2}t\sin(\frac{\pi}{2}u))^{t}+t(t-1)\frac{\pi}{2}\cos(\frac{\pi}{2}u))^{t-2}\sin^{2}(\frac{\pi}{2})-\frac{\pi}{2}t\cos(\frac{\pi}{2}u))^{t}],\\
    \quad g^{''}(0) =-\frac{\pi^3}{8}t.
$$
We take $T_1 = \frac{\pi}2 -\frac{\pi^{3}t}{16}u^{2}$.  At this point, we define $Y$ in terms of $T_1$, but we need to have mean $0$. Therefore, we compute:
$$
    \mathbb{E}(T_1)=\frac{\pi}2 -\frac{\pi^{3}t}{16}\mathbb{E}(u^{2})=\frac{\pi}2 -\frac{\pi^{3}}{16}t*\frac{1}{3}=\frac{\pi}2 -\frac{\pi^{3}}{48}t,
$$
then we choose $Y=T_1-(\frac{\pi}2 -\frac{\pi^{3}}{48}t)=\frac{\pi}2 -\frac{\pi^{3}t}{16}u^{2} -\frac{\pi}2 + \frac{\pi^{3}}{48}t = \frac{\pi^3}{48}t-(\frac{\pi^3}{16}t)u^{2}$. We estimate $\theta$ via the following algorithm:

\begin{itemize}
  \item Simulate $U_1, U_2, \ldots, U_n \overset{d}{\sim} \mathcal{U}(0,1)$.
  \item Simulate $X$: $g(u_1), g(u_2), \ldots, g(u_n)$.
  \item Simulate $Y$: $\frac{\pi^3}{48}t-(\frac{\pi^3}{16}t)u_1^{2}, \frac{\pi^3}{48}t-(\frac{\pi^3}{16}t)u_2^{2}, \ldots, \frac{\pi^3}{48}t-(\frac{\pi^3}{16}t)u_n^{2}$.
  \item Calculate the $\text{Cov}(X,Y)$ and $\text{Var}(Y)$ with the samples to determine $c^{\star}$.
  \item Approximate $\theta$ with the mean of $X+c^{\star}Y$ and standard errors with the standard errors of $X+c^{\star}Y$.
\end{itemize}

##### Using for loop to investigate the performance of integral across different values of t:
###### CV1
```{r}
set.seed(1)
EstIntegral_cv1 <- numeric(length(grid_t))
Diff_cv1 <- numeric(length(grid_t))
SE_cv1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  X <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  Y <- pi^3/48*t-pi^3/16*t*u^2
  
  c <- -cov(X,Y)/var(Y)
  X_cv1 <- X+c*Y
  EstIntegral_cv1[t] <- mean(X_cv1)
  Diff_cv1[t] <- exact_values_C[t] -  mean(X_cv1)
  SE_cv1[t] <- sd(X_cv1)/sqrt(n)
}

#cbind(Mean_diff_cv=Diff_cv1, SE_cv = mean(SE_cv1))
Diff_cv1
print('-----------------------------')
SE_cv1
print('-----------------------------')
Std_Error_3
mean(Std_Error_3)
```

###### CV2
```{r}
set.seed(1)
EstIntegral_cv2 <- numeric(length(grid_t))
Diff_cv2 <- numeric(length(grid_t))
SE_cv2 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  X_tilde <- ((sin(pi/2*u + pi/4))^t) * pi/2
  Y_2 <- X_tilde - mean(X_tilde)
  c <- -cov(X,Y_2)/var(Y_2)
  
  theta_hat <- X+c*Y_2
  EstIntegral_cv2[t] <- mean(theta_hat)
  SE_cv2[t] <- sd(theta_hat)/sqrt(n)
  Diff_cv2[t] <- exact_values_C[t] - mean(theta_hat)
}



# cbind(mean = mean_cv2, Mean_diff_cv=Diff_cv2, SE_cv = mean(SE_cv2))
# improve_1 <- abs(mean(SE_cv2)-mean(Std_Error_3))/mean(Std_Error_3)
# improve_1
# cat("The control variate method reduced the standard error by",improve_1)

Diff_cv2
print('-----------------------------')
SE_cv2
print('-----------------------------')
Std_Error_3
mean(SE_cv2)
```



###### CV3
```{r}
set.seed(1)
EstIntegral_cv3 <- numeric(length(grid_t))
Diff_cv3 <- numeric(length(grid_t))
SE_cv3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  X <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  X_tilde <- ((pi/2*u - (pi/2*u)^3/6)^t + (1 - (pi/2*u)^2/2)^t) * pi/2
  Y <- X_tilde - mean(X_tilde)
  
  c <- -cov(X,Y)/var(Y)
  X_cv3 <- X+c*Y
  EstIntegral_cv3[t] <- mean(X_cv3)
  Diff_cv3[t] <- exact_values_C[t] -  mean(X_cv3)
  SE_cv3[t] <- sd(X_cv3)/sqrt(n)
}

```

```{r}
round((Std_Error_3 - SE_cv3)/Std_Error_3 * 100, digits = 2)
```
```{r}
Diff_cv3
```

#### Stratified Sampling
```{r}
# SS
set.seed(1)

#X_ss <- matrix(numeric(n*length(grid_t)), nrow = length(grid_t))
EstIntegral_ss <- numeric(length(grid_t))
Diff_ss <- numeric(length(grid_t))
Std_Error_ss <- numeric(length(grid_t))

for (t in grid_t){
  j <- 1:n  # split the integral [0, 1] into 10^5 smaller intervals
  u <- runif(n)
  
  sample_ss <- (sin(pi/2*(u + j - 1)/n)^t + cos(pi/2*((u + j - 1)/n))^t) * pi/2
  #X_ss[t, ] <- sample_ss
  EstIntegral_ss[t] <- mean(sample_ss)
  Diff_ss[t] <- exact_formula_C(t) - EstIntegral_ss[t]
  Std_Error_ss[t] <- sd(sample_ss)/sqrt(n)
}
```


#### Stratified Sampling + Control Variates
```{r}
## SS + CV
EstIntegral_cvss <- numeric(length(grid_t))
Diff_cvss <- numeric(length(grid_t))
SE_cvss <- numeric(length(grid_t))

for (t in grid_t){
  j <- 1:n  # split the integral [0, 1] into 10^5 smaller intervals
  u <- runif(n)

  X <- (sin(pi/2*(u + j - 1)/n)^t + cos(pi/2*(u + j - 1)/n)^t) * pi/2
  X_tilde <- ((pi/2*(u + j - 1)/n - (pi/2*(u + j - 1)/n)^3/6)^t + (1 - (pi/2*(u + j - 1)/n)^2/2)^t) * pi/2
  Y <- X_tilde - mean(X_tilde)
  
  c <- -cov(X,Y)/var(Y)
  X_cvss <- X+c*Y
  EstIntegral_cvss[t] <- mean(X_cvss)
  Diff_cvss[t] <- exact_values_C[t] -  mean(X_cvss)
  SE_cvss[t] <- sd(X_cvss)/sqrt(n)
}

```

```{r}
# Compare SS + CV with SS
abs(Diff_ss) - abs(Diff_cvss)
round((Std_Error_ss - SE_cvss)/Std_Error_ss, digits = 2)

# Compare SS + CV with CV
abs(Diff_cv3) - abs(Diff_cvss)
round((SE_cv3 - SE_cvss)/SE_cv3, digits = 2)

# Compare SS + CV with pure Monte Carlo without any variance reduction
abs(Diff_3) - abs(Diff_cvss)
round((Std_Error_3 - SE_cvss)/Std_Error_3, digits = 2)

# Compare CV with pure Monte Carlo without any variance reduction
abs(Diff_3) - abs(Diff_cv3)
round((Std_Error_3 - SE_cv3)/Std_Error_3, digits = 2)
```


```{r}
# set.seed(1)
# n <- 10^6
# u <- runif(n)
# j <- 1:n
# sample_1a2 <- ((sin(pi/2*((u+j-1)/n)))^t + (cos(pi/2*((u+j-1)/n)))^t) * pi/2
# 
# Mu_ss1 <- mean(sample_1a2)
# Diff_ss1 <- EstIntegral_3 - Mu_ss1
# SE_ss1 <- sd(sample_1a2)/sqrt(n)
# 
# cbind(Mu_ss1, Diff_ss1, SE_ss1)
# improve_SS <- abs(Mu_ss1-EstIntegral_3)/EstIntegral_3
```


The following table shows the comparison between different variance reduction methods,
```{r}
df_vr1<-data.frame(
  Method=c("Transformation 3","VC_1a","SS_1a"),
  Mean=c(mean(sample_3),mean(EstIntegral_2),mean(EstIntegral_3)),
  Mean_diff=c(mean(Diff_3),mean(Diff_cv2),mean(Diff_ss1)),
  Standar_error=c(mean(Std_Error_3),mean(SE_cv2),mean(SE_ss1))
)
df_vr1
```



## (b)
### Calculate the true value of the double integral
```{r echo=TRUE}
set.seed(1)
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
true_integral_1b <- result$value
true_integral_1b
```

#### Alternatively, using another package:
```{r}
set.seed(1)
# install.packages('pracma')
library(pracma)

f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r) 
}

dblquad(f, -1, 1, -1, 1)
```

### Pure Monte Carlo Methods
Since the equation os symmetric, we can compute: 
$$
    \int_{-1}^{1}\int_{-1}^{1}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy{=}4\int_{0}^{1}\int_{0}^{1}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy  
    \overset{(X, Y) ind \overset{d}{\sim} \mathcal{U}(0,1) }{=}\\
    4\mathbb{E}[\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}} ] \overset{MC}{\approx} \frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy ,
$$
where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.

```{r}
set.seed(1)

n <- 10^5
x <- runif(n)
y <- runif(n)
sample_mc <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)

EstIntegral_mc <- mean(sample_mc)
Diff_mc <- true_integral_1b - EstIntegral_mc
Std_Error_mc <- sd(sample_mc)/sqrt(n)

rbind(EstIntegral_mc, Diff_mc, Std_Error_mc)
```

#### 1.Antithetic Variate Reduction
We need to calculate
$$
\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(x^2 + y^2)})}{\sqrt{x^2 + y^2}}=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{((1-x)^2 + (1-y)^2)})}{\sqrt{(1-x)^2 + (1-y)^2}},
$$
where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.
```{r}
set.seed(1)
n <- 10^5
x <- runif(n)
y <- runif(n)
sample1 <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)
sample2 <- 4 * sin(sqrt((1-x)^2 + (1-y)^2))/sqrt((1-x)^2 + (1-y)^2)
EstIntegral_av <- mean((sample1+sample2)/2)
Mean_diff_av <- true_integral_1b - EstIntegral_av
Std_Error_av <- sd((sample1+sample2)/2)/sqrt(n)

rbind(EstIntegral_av, Mean_diff_av, Std_Error_av)
improve_av <- abs(Std_Error_av-Std_Error_mc)/Std_Error_mc
improve_av
```
The antithetic variate reduction reduces the standard error by 75.8793%.

#### 2.Stratified Sampling
Since $1 - U_i \overset{d}{=} U_i$, we can compute
$$
h(U)=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(\frac{(U_i+i-1)}{n})^{2} + ((\frac{(U_j+j-1)}{n})^2})}{\sqrt{(\frac{(U_i+i-1)}{n})^{2} + ((\frac{(U_j+j-1)}{n})^2})},
$$
where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.
```{r}
set.seed(1)
n <- 10^5
u <- runif(n)
j <- 1:n
sample_1b2 <- 4 * (sin(sqrt(((u+j-1)/n)^2 + ((u+j-1)/n)^2)))/sqrt(((u+j-1)/n)^2 + ((u+j-1)/n)^2)

Mu_SS <- mean(sample_1b2)
Diff_SS <- true_integral_1b - Mu_SS
SE_SS <- sd(sample_1b2)/sqrt(n)

rbind(Mu_SS, Diff_SS, SE_SS)
improve_SS <- abs(Mu_SS-EstIntegral_mc)/EstIntegral_mc
improve_SS
```

The following table shows the comparison between different methods:
```{r}
df_1b<-data.frame(
  Method=c("Ture Value_1b","MC_1b","AV_1b","SS_1b"),
  Mean=c(true_integral_1b,mean(EstIntegral_mc),mean(EstIntegral_av),mean(Mu_SS)),
  Mean_diff=c("NA",mean(Diff_mc),mean(Mean_diff_av),mean(Diff_SS)),
  Standar_error=c("NA",mean(Std_Error_mc),mean(Std_Error_av),mean(SE_SS))
)
df_1b
```

Therefore, the antithetic variate reduction can improve the accuarcy by 75.8793%, which is the best.