---
title: "Q1"
output:
  pdf_document: default
  html_document: default
date: "2025-04-07"
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a)
### The true values
We compute the exact values of C(t) for t values from 1 to 100.
```{r include=FALSE}
grid_t <- 1:100

exact_formula_C <- function(t){
  (sqrt(pi) * gamma((t + 1)/2))/gamma(t/2 + 1)
}

exact_values_C <- exact_formula_C(grid_t)
```

### Monte Carlo Methods
We take n=10^6.
```{r include=FALSE}
n <- 10^6
```

#### Transformation 1

For transformation 1, we use $u=sinx$, then $x=arcsin(u)$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{x=\arcsin(u)}{=} \int_{0}^{1} {u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[{u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}}\right] \overset{MC}{\approx} \sum_{i=1}^{N}[u^{t}[\cos(\arcsin(u)]^{-1}+[\cos(\arcsin(u)]^{t-1}]
$$

##### Using for loop to investigate the performance of integral across different values of t:
```{r include=FALSE}
set.seed(1)
EstIntegral_1 <- numeric(length(grid_t))
Diff_1 <- numeric(length(grid_t))
Std_Error_1 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_1 <- u^t*1/(cos(asin(u)))+(cos(asin(u)))^(t-1)
  
  EstIntegral_1[t] <- mean(sample_1)
  Diff_1[t] <- exact_formula_C(t) - EstIntegral_1[t]
  Std_Error_1[t] <- sd(sample_1)/sqrt(n)
}

rbind(mean_1=mean(sample_1),mean_diff_1 = mean(Diff_1), mean_SE_1 = mean(Std_Error_1))
```

#### Transformation 2
For transformation 2, we use $u=cosx$, then $x=\arccos(u)$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{ x=\arccos(u)}{=} \int_{0}^{1} {u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[{u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}}\right] \overset{MC}{\approx} \sum_{i=1}^{N}[u^{t}[\sin(\arccos(u)]^{-1}+[\sin(\arccos(u)]^{t-1}]
$$
##### Using for loop to investigate the performance of integral across different values of t:
```{r include=FALSE}
set.seed(1)
EstIntegral_2 <- numeric(length(grid_t))
Diff_2 <- numeric(length(grid_t))
Std_Error_2 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_2 <- (sin(acos(u)))^{t-1} + u^t*(sin(acos(u)))^{-1}
  
  EstIntegral_2[t] <- mean(sample_2)
  Diff_2[t] <- exact_formula_C(t) - EstIntegral_2[t]
  Std_Error_2[t] <- sd(sample_2)/sqrt(n)
}

rbind(mean_2=mean(sample_2),mean_diff_2 = mean(Diff_2), mean_SE_2 = mean(Std_Error_2))
```
According to the above results, transformation 1 and 2 have exactly the same mean and standard error.

#### Transformation 3
For transformation 3, we use $u=\frac{2}\pi*x$, then $x=\frac{\pi}2u$. When $x=0$, then $u=0$, and when $x=\pi/2$, then $u=1$.
We compute:
$$
\int_{0}^{\pi/2} {(\sin(x))^{t}+(\cos(x))^{t}}\,dx \overset{x=\frac{\pi}2u}{=} \int_{0}^{1} {[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]\frac{\pi}2}\,du  \overset{\mathcal{U}\overset{d}{=}\mathcal{U}(0,1)} {=}\\ \mathbb{E}\left[\frac{\pi}2[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]\right] \overset{MC}{\approx} \sum_{i=1}^{N}\frac{\pi}2[(\sin((\frac{\pi}2u))^{t}+((\cos(\frac{\pi}2u))^{t}]
$$
##### Using for loop to investigate the performance of integral across different values of t:
```{r include=FALSE}
set.seed(1)
EstIntegral_3 <- numeric(length(grid_t))
Diff_3 <- numeric(length(grid_t))
Std_Error_3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  sample_3 <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  
  EstIntegral_3[t] <- mean(sample_3)
  Diff_3[t] <- exact_formula_C(t) - EstIntegral_3[t]
  Std_Error_3[t] <- sd(sample_3)/sqrt(n)
}

rbind(mean_diff_3 = mean(Diff_3), SE_3 = mean(Std_Error_3))
```

The following table shows the comparison between different methods:
```{r echo=FALSE}
df_ts<-data.frame(
  Transformation = c("Transformation 1","Transformation 2","Transformation 3"),
  Mean_Diff=c(mean(Diff_1),mean(Diff_2),mean(Diff_3)),
  Mean_Standard_Error=c(mean(Std_Error_1),mean(Std_Error_2),mean(Std_Error_3))
)
df_ts
```

According to the transformation methods above, transformation 3 has the smallest difference between estimated mean and actual mean, and also the smallest standard error. Thus we do the variance reduction based on transformation 3.

The antithetic variate reduction is not applicable since the function is not monotone. Then we try control variates method first.

#### 1.Control variates method

We want to find a random variable $Y$ with mean $0$ and a good constant $c$ such that $\mathbb{E}(X+cY) \approx \theta$.

\begin{itemize}
  \item Simulate $U_1, U_2, \ldots, U_n \overset{d}{\sim} \mathcal{U}(0,1)$.
  \item Simulate $\tilde{X}=(\frac{\pi}{2}u-\frac{(\frac{\pi}{2}u)^{3}}{6})^{t}+(1-\frac{(\frac{\pi}{2}u)^{2}}{2})^{t}$, where $(\frac{\pi}{2}u-\frac{(\frac{\pi}{2}u)^{3}}{6})$ is the Taylor expansion of $\sin(\frac{\pi}{2}u)$, and $(1-\frac{(\frac{\pi}{2}u)^{2}}{2})$ is the Taylor expansion of $\cos(\frac{\pi}{2}u)$.
  \item Simulate $Y=\tilde{X}-\mathbb{E}\left(\tilde{X}\right)$.
  \item Calculate the $\text{Cov}(X,Y)$ and $\text{Var}(Y)$ with the samples to determine $c^{\star}$, $c^{\star}=-\frac{\text{Cov}(X,Y)}{\text{Var}(Y)}$
  \item Approximate $\theta$ with the mean of $X+c^{\star}Y$ and standard errors with the standard errors of $X+c^{\star}Y$.
\end{itemize}

###### Control Variate Method
```{r include=FALSE}
set.seed(1)
EstIntegral_cv3 <- numeric(length(grid_t))
Diff_cv3 <- numeric(length(grid_t))
SE_cv3 <- numeric(length(grid_t))

for (t in grid_t){
  u <- runif(n)
  X <- ((sin(pi/2*u))^t + (cos(pi/2*u))^t) * pi/2
  X_tilde <- ((pi/2*u - (pi/2*u)^3/6)^t + (1 - (pi/2*u)^2/2)^t) * pi/2
  Y <- X_tilde - mean(X_tilde)
  
  c <- -cov(X,Y)/var(Y)
  X_cv3 <- X+c*Y
  EstIntegral_cv3[t] <- mean(X_cv3)
  Diff_cv3[t] <- exact_values_C[t] -  mean(X_cv3)
  SE_cv3[t] <- sd(X_cv3)/sqrt(n)
}
```

#### Stratified Sampling
We estimate:
$$
h(U)=[(\sin(\frac{\pi}{2}\frac{U_j + j - 1}{n}))^{t} + (\cos(\frac{\pi}{2}\frac{U_j + j - 1}{n}))^t]\frac{\pi}{2}
$$

```{r include=FALSE}
set.seed(1)

EstIntegral_ss <- numeric(length(grid_t))
Diff_ss <- numeric(length(grid_t))
Std_Error_ss <- numeric(length(grid_t))

for (t in grid_t){
  j <- 1:n  # split the integral [0, 1] into 10^5 smaller intervals
  u <- runif(n)
  
  sample_ss <- (sin(pi/2*(u + j - 1)/n)^t + cos(pi/2*((u + j - 1)/n))^t) * pi/2
  EstIntegral_ss[t] <- mean(sample_ss)
  Diff_ss[t] <- exact_formula_C(t) - EstIntegral_ss[t]
  Std_Error_ss[t] <- sd(sample_ss)/sqrt(n)
}
```


#### Stratified Sampling + Control Variates
```{r include=FALSE}
EstIntegral_cvss <- numeric(length(grid_t))
Diff_cvss <- numeric(length(grid_t))
SE_cvss <- numeric(length(grid_t))

for (t in grid_t){
  j <- 1:n  # split the integral [0, 1] into 10^5 smaller intervals
  u <- runif(n)

  X <- (sin(pi/2*(u + j - 1)/n)^t + cos(pi/2*(u + j - 1)/n)^t) * pi/2
  X_tilde <- ((pi/2*(u + j - 1)/n - (pi/2*(u + j - 1)/n)^3/6)^t + (1 - (pi/2*(u + j - 1)/n)^2/2)^t) * pi/2
  Y <- X_tilde - mean(X_tilde)
  
  c <- -cov(X,Y)/var(Y)
  X_cvss <- X+c*Y
  EstIntegral_cvss[t] <- mean(X_cvss)
  Diff_cvss[t] <- exact_values_C[t] -  mean(X_cvss)
  SE_cvss[t] <- sd(X_cvss)/sqrt(n)
}

```

#### Compare all variance reduction methods
```{r}
# Compare SS + CV with SS
abs(Diff_ss) - abs(Diff_cvss)
round((Std_Error_ss - SE_cvss)/Std_Error_ss, digits = 2)

# Compare SS + CV with CV
abs(Diff_cv3) - abs(Diff_cvss)
round((SE_cv3 - SE_cvss)/SE_cv3, digits = 2)

# Compare SS + CV with pure Monte Carlo without any variance reduction
abs(Diff_3) - abs(Diff_cvss)
round((Std_Error_3 - SE_cvss)/Std_Error_3, digits = 2)

# Compare CV with pure Monte Carlo without any variance reduction
abs(Diff_3) - abs(Diff_cv3)
round((Std_Error_3 - SE_cv3)/Std_Error_3, digits = 2)
```

## (b)
### Calculate the true value of the double integral
```{r echo=TRUE}
set.seed(1)
f <- function(x, y) {
  r <- sqrt(x^2 + y^2)
  ifelse(r == 0, 1, sin(r) / r)  # Handle the removable singularity at (0,0)
}

result <- integrate(function(y) {
  sapply(y, function(yi) {
    integrate(function(x) f(x, yi), -1, 1)$value
  })
}, -1, 1)

# Print the result
true_integral_1b <- result$value
true_integral_1b
```

### Pure Monte Carlo Methods
Since the equation is symmetric, we can compute: 
$$
    \int_{-1}^{1}\int_{-1}^{1}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy{=}4\int_{0}^{1}\int_{0}^{1}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy  
    \overset{(X, Y) ind \overset{d}{\sim} \mathcal{U}(0,1) }{=}\\
    4\mathbb{E}[\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}} ] \overset{MC}{\approx} \frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{x^{2}+y^{2}})}{\sqrt{x^{2}+y^{2}}}\,dx\,dy ,
$$
where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.

```{r include=FALSE}
set.seed(1)

n <- 10^5
x <- runif(n)
y <- runif(n)
sample_mc <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)

EstIntegral_mc <- mean(sample_mc)
Diff_mc <- true_integral_1b - EstIntegral_mc
Std_Error_mc <- sd(sample_mc)/sqrt(n)

rbind(EstIntegral_mc, Diff_mc, Std_Error_mc)
```

#### 1.Antithetic Variate Reduction
We need to calculate
$$
\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(x^2 + y^2)})}{\sqrt{x^2 + y^2}}=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{((1-x)^2 + (1-y)^2)})}{\sqrt{(1-x)^2 + (1-y)^2}},
$$
where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.
```{r include=FALSE}
set.seed(1)
n <- 10^5
x <- runif(n)
y <- runif(n)
sample1 <- 4 * sin(sqrt(x^2 + y^2))/sqrt(x^2 + y^2)
sample2 <- 4 * sin(sqrt((1-x)^2 + (1-y)^2))/sqrt((1-x)^2 + (1-y)^2)
EstIntegral_av <- mean((sample1+sample2)/2)
Mean_diff_av <- true_integral_1b - EstIntegral_av
Std_Error_av <- sd((sample1+sample2)/2)/sqrt(n)

rbind(EstIntegral_av, Mean_diff_av, Std_Error_av)
```

```{r}
improve_av <- (Std_Error_mc - Std_Error_av)/Std_Error_mc
improve_av
```

The antithetic variate reduction reduces the standard error by 75.8793%.

#### 2.Stratified Sampling
Since $1 - U_i \overset{d}{=} U_i$, we can compute
$$
h(U)=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2})}{\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2}},
$$
where $X_1,\ldots,X_N \overset{d}{\sim} \mathcal{U}(0,1)$, $Y_1,\ldots,Y_N \overset{d}{\sim} \mathcal{U}(0,1)$, and the samples are independent.
```{r include=FALSE}
set.seed(1)
n <- 10^5
j <- 1:n
sample_1b2 <- 4 * (sin(sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)))/sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)

Mu_SS <- mean(sample_1b2)
Diff_SS <- true_integral_1b - Mu_SS
SE_SS <- sd(sample_1b2)/sqrt(n)

rbind(Mu_SS, Diff_SS, SE_SS)
```

```{r}
improve_SS <- (EstIntegral_mc - Mu_SS)/EstIntegral_mc
improve_SS
```


#### Stratified Sampling + Antithetic Variate reduction
To combine the stratified sampling with antithetic variate reduction, we calculte the mean of the following to estimations:
$$
h(U)=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2}}{\sqrt{(\frac{U_i+i-1}{n})^{2} + (\frac{U_j+j-1}{n})^2}},
$$
and
$$
h(1-U)=\frac{4}{N}\sum_{i=1}^{N}\frac{\sin(\sqrt{(\frac{i-U_i}{n})^{2} + (\frac{j-U_j}{n})^2}}{\sqrt{(\frac{i-U_i}{n})^{2} + (\frac{j-U_j}{n})^2}}.
$$

```{r include=FALSE}
set.seed(1)
n <- 10^5
j <- 1:n
sample_avss_1 <- 4 * (sin(sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)))/sqrt(((x+j-1)/n)^2 + ((y+j-1)/n)^2)
sample_avss_2 <- 4 * (sin(sqrt(((j-x)/n)^2 + ((j-y)/n)^2)))/sqrt(((j-x)/n)^2 + ((j-y)/n)^2)
sample_avss <- (sample_avss_1 + sample_avss_2)/2

Mu_avss <- mean(sample_avss)
Diff_avss <- true_integral_1b - Mu_avss
SE_avss <- sd(sample_avss)/sqrt(n)

rbind(Mu_avss, Diff_avss, SE_avss)
```

```{r}
improve_SS <- (EstIntegral_mc - Mu_avss)/EstIntegral_mc
improve_SS
```

The following table shows the comparison between different methods:
```{r echo=FALSE}
df_1b<-data.frame(
  Method=c("Ture Value_1b","MC_1b","AV_1b","SS_1b","AVSS_1b"),
  Mean=c(true_integral_1b,EstIntegral_mc,EstIntegral_av,Mu_SS,Mu_avss),
  Diff=c("NA",Diff_mc,Mean_diff_av,Diff_SS,Diff_avss),
  Standar_error=c("NA",Std_Error_mc,Std_Error_av,SE_SS,SE_avss)
)
df_1b
```
Therefore, the antithetic variate reduction can reduce the variance by 75.8793%, and the difference between estimated mean and the true mean is the smallest. Therefore, the antithetic variate reduction is the best method to improve the accuracy.