---
title: "Q2"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```

# Question 2
## a)
### Generate various values of m, nu, lambda and alpha
```{r}
m_list <- seq(1.6, 10, length.out = 10)
nu_list <- seq(-5, 5, length.out = 10)
lambda_list <- seq(-5, 5, length.out = 10)
alpha_list <- seq(0.1, 10, length.out = 10)

groups <- expand.grid(m_list, nu_list, lambda_list, alpha_list)

m <- groups$Var1
nu <- groups$Var2
lambda <- groups$Var3
alpha <- groups$Var4
```


```{r}
theoretical_mean <- lambda - alpha*nu/(2*(m - 1))
theoretical_var <- alpha^2/(2*m - 3) * (1 + nu^2/(4*(m - 1)^2))
```


```{r}
Q2_summary <- data.frame(m, nu, lambda, alpha, theoretical_mean, theoretical_var)
```



### With transformation
$$
y = tan^{-1}(\frac{x - \lambda}{\alpha}), y \in [-\frac{\pi}{2}, \frac{\pi}{2}]
$$
We get
$$
f(y) \propto (cos(y))^{2m}e^{-\nu y}
$$


#### Envelope function 3 (cosine)
$$
g(y) = \frac{1}{2}cos(y)
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = sin^{-1}(2U - 1)
$$

```{r}
# Generate Y
n <- 10^5
U <- runif(n)

Y <- asin(2*U-1)
```

```{r}
# Generate the sample using AR
ptm <- proc.time()
U <- runif(n)

temp <- atan(-nu/(2*m-1))

sample_envelope_3 <- lapply(1:length(nu), function(i){
  temp <- temp[i]
  m <- m[i]
  nu <- nu[i]
  
  Y[U < (cos(Y)^{2*m-1}*exp(-nu*Y))/(cos(temp)^{2*m-1}*exp(-nu*temp))]
})
sim_envelope_3 <- (proc.time()- ptm)[3]

plot(density(sample_envelope_3[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_3 <- lapply(1:length(nu), function(i) tan(sample_envelope_3[[i]]) * alpha[i] + lambda[i])

envelop_3_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_3[[i]])))
envelop_3_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_3[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_3_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_3[[i]])/10^5))
```

```{r}
# update Q2_summary
Q2_summary <- Q2_summary %>% 
  mutate(envelop_3_empirical_mean, envelop_3_empirical_var, envelope_3_empirical_acceptance)
Q2_summary
```

## c)
### Empirical quantiles
```{r}
quantile_percents <- c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)

matrix <- matrix(numeric(length = length(nu)*length(quantile_percents)), nrow = length(nu))

for (i in 1:length(nu)){
  matrix[i,] <- quantile(X_envelope_3[[i]], prob = quantile_percents)
}
```

```{r}
Q2_empirical_quantiles <- as.data.frame(matrix) %>% mutate(index = 1:length(nu))
colnames(Q2_empirical_quantiles) <- c(paste0('Empirical_', quantile_percents*100, '%'), 'index')

Q2_summary_quantiles <- Q2_summary %>% 
  mutate(index = 1:length(nu)) %>%
  select(m, nu, lambda, alpha, index, theoretical_mean, theoretical_var) %>%
  left_join(Q2_empirical_quantiles, by = join_by(index)) %>%
  select(-index)
Q2_summary_quantiles
```


### CLT quantiles
Try to make the unrealistic assumption of 
$$
\frac{X - E(X)}{\sqrt{Var(X)}} \sim N(0, 1)
$$

and estimate quantiles accordingly:
$$
quantile = E(X) + \sqrt{Var(X)}\Phi^{-1}(p)
$$

```{r}
matrix <- matrix(numeric(length = length(nu)*length(quantile_percents)), nrow = length(nu))

for (i in 1:length(nu)){
  matrix[i,] <- Q2_summary$theoretical_mean[i] + sqrt(Q2_summary$theoretical_var[i]) * qnorm(quantile_percents, mean = 0, sd = 1)
}
```

```{r}
Q2_CLT_quantiles <- as.data.frame(matrix) %>% mutate(index = 1:length(nu))
colnames(Q2_CLT_quantiles) <- c(paste0('CLT_', quantile_percents*100, '%'), 'index')

Q2_summary_quantiles <- Q2_summary_quantiles %>% 
  mutate(index = 1:length(nu)) %>%
  left_join(Q2_CLT_quantiles, by = join_by(index)) %>%
  select(-index)
Q2_summary_quantiles
```

### Compare two types of quantiles:
```{r}
Q2_summary_quantiles_diff <- Q2_summary_quantiles %>%
  mutate(`0.05` = `CLT_5%` - `Empirical_5%`,
         `0.10` = `CLT_10%` - `Empirical_10%`,
         `0.25` = `CLT_25%` - `Empirical_25%`,
         `0.50` = `CLT_25%` - `Empirical_25%`,
         `0.75` = `CLT_75%` - `Empirical_75%`,
         `0.90` = `CLT_90%` - `Empirical_90%`,
         `0.95` = `CLT_95%` - `Empirical_95%`) %>%
  select(m, nu, lambda, alpha, `0.05`, `0.10`, `0.25`, `0.50`, `0.75`, `0.90`, `0.95`)

Q2_summary_quantiles_diff
```



```{r}
temp <- Q2_summary_quantiles_diff %>%
  mutate(m = round(m, digits = 3), nu = round(nu, digits = 3)) %>%
  filter(!m %in% c(1.6, 2.533)) %>% # Remove the two columns to make plots nicer
  select(m, nu, `0.05`, `0.10`, `0.25`, `0.50`, `0.75`, `0.90`, `0.95`)

temp_long <- pivot_longer(
 temp,
 cols = c(`0.05`, `0.10`, `0.25`, `0.50`, `0.75`, `0.90`, `0.95`),
 names_to = 'x',
 values_to = "y"
)

ggplot(temp_long, aes(x = x, y = y, color = factor(m))) +
  geom_boxplot() +
  facet_wrap(~nu, labeller = label_both) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Percents", y = "CLT Percentile - Sample Percentile", color = "m", title = 'Difference between Sample Percentiles and CLT Percentiles, holding nu constant') + 
  scale_x_discrete(labels = paste0(quantile_percents*100, '%'))

ggplot(temp_long, aes(x = x, y = y, color = factor(nu))) +
  geom_boxplot() +
  facet_wrap(~m, labeller = label_both) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Percents", y = "CLT Percentile - Sample Percentile", color = "nu", title = 'Difference between Sample Percentiles and CLT Percentiles, holding m constant') +
  scale_x_discrete(labels = paste0(quantile_percents*100, '%'))
```

Conclusions:
CLT percentiles are lower than the sample percentiles when the percentage is low.
CLT percentiles are higher than the sample percentiles when the percentage is high.
Therefore, we can conclude that compared to a normal distribution, our sample distribution has heavier tails on both sides.
We also notice the following facts:
- The higher the $m$, the lower the discrepancies between the CLT percentiles and the sample percentiles. Hence, as $m$ gets higher, the sample distribution converges to a normal distribution.
- The closer $\nu$ is to zero, the lower discrepancies between the CLT percentiles and the sample percentiles. Hence, as $\nu$ gets closer to zero, the sample distribution coverges to a normal distribution.



