---
title: "Q2"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

## Question 2 
### (a)\
We generate 10 values for each of the parameters $m$, $\nu$, $\lambda$, $\alpha$, based on the conditions on their values specified in the question. They are shuffled so that $10^4 = 10000$ distinct sets of parameters are created. \
```{r}
m_list <- seq(1.6, 10, length.out = 10)
nu_list <- seq(-5, 5, length.out = 10)
lambda_list <- seq(-5, 5, length.out = 10)
alpha_list <- seq(0.1, 10, length.out = 10)

groups <- expand.grid(m_list, nu_list, lambda_list, alpha_list)

m <- groups$Var1
nu <- groups$Var2
lambda <- groups$Var3
alpha <- groups$Var4
```

We also calculate the theoretical mean and variance for each set of parameters using the formula from part (b): \
```{r}
theoretical_mean <- lambda - alpha*nu/(2*(m - 1))
theoretical_var <- alpha^2/(2*m - 3) * (1 + nu^2/(4*(m - 1)^2))
```

Next, we create a dataframe summarising all information about the future simulated datasets. This dataset has 10000 rows, each row representing the information of one set of parameters generated before. \
```{r}
Q2_summary <- data.frame(m, nu, lambda, alpha, 
                         theoretical_mean, theoretical_var)
head(Q2_summary)
```


With transformation \
$$
y = \tan^{-1}(\frac{x - \lambda}{\alpha}), y \in [-\frac{\pi}{2}, \frac{\pi}{2}]
$$
We get \
$$
f(y) \propto (\cos(y))^{2m}e^{-\nu y}
$$
and we are able to derive three envelop functions from that: \
**Envelope function 1** - this is essentially a uniform distribution \
$$
g(y)=\frac{1}{\pi}
$$

**Envelope function 2** \
$$
g(y)\propto e^{-\nu y}
$$
**Envelope function 3** \
$$
g(y)\propto\cos(y)
$$

We can derive two more envelope functions from the original form of the target function, i.e.
$$
f(x) \propto \left[ 1 + (\frac{x - \lambda}{\alpha})\right]^{-m} e^{-\nu\ arctan(\frac{x - \lambda}{\alpha})}
$$

**Envelope function 4** - after writing it down, we noticed that this is essentially a Cauchy distribution \
$$
g(y) \propto (1+(\frac{y-\lambda}{\alpha})^{2})^{-1}
$$
**Envelope function 5**
$$
g(y) \propto (1+(\frac{y-\lambda}{\alpha})^{2})^{-m}
$$
By setting $m = \frac{k+1}{2}$ and $z = \frac{\sqrt{k}(y - \lambda)}{\alpha}$, we transform it to a form similar to a t-distribution: \
$$
g(y) \propto (1+\frac{z^{2}}{k})^{-\frac{k+1}{2}}
$$

We run R codes across all sets of parameters to get the **sample mean**, **sample variance** and **empirical acceptance rate** for each envelope function and each set of parameters. These numbers will be integrated into `Q2_summary` dataset as new columns so that we can compare the performance of different envelope functions later.\

Unfortunately, due to the limit on number of pages, we are unable to explain every algorithm we wrote in R. We will elaborate on the best envelope function we picked later on, after the best envelope function is picked.\
R will take approximately 12 minutes to run the codes for all 5 envelope functions. That is a reasonable efficiency for 10000 sets of parameters. \

#### Codes for Envelope function 1 (uniform)
```{r include=FALSE}
# Generate Y
set.seed(1)
n <- 10^5
Y <- runif(n) * pi - pi/2
```

```{r include=FALSE}
# Generate the sample using AR
set.seed(1)
ptm <- proc.time()
U <- runif(n)

sample_envelope_1 <- lapply(1:length(nu), function(i) {
  m <- m[i]
  nu <- nu[i]
  
  temp <- atan(-nu / (2 * m))
  accepted_Y <- Y[U < 
    (cos(Y))^(2 * m) * exp(-nu * Y) / (cos(temp)^(2 * m) * exp(-nu * temp))]
})
sim_envelope_1 <- (proc.time()- ptm)[3]

plot(density(sample_envelope_1[[1]]), type = 'l')
```

```{r include=FALSE}
# Produce X, and check mean and variance
X_envelope_1 <- lapply(1:length(nu), function(i){tan(sample_envelope_1[[i]]) * alpha[i] + lambda[i]})
# X_envelope_1[[1]][1:10]
# sample_envelope_1[[1]][1:10]

envelop_1_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_1[[i]])))
envelop_1_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_1[[i]])))
```


```{r include=FALSE}
# empirical rate of acceptance
envelope_1_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_1[[i]])/10^5))
```

```{r include=FALSE}
# add the results to the summary
library(tidyverse)
Q2_summary <- Q2_summary %>% 
  mutate(envelop_1_empirical_mean, envelop_1_empirical_var, envelope_1_empirical_acceptance)

Q2_summary
```

#### Codes for Envelope function 2 (exponential)
Envelope function 2 is not a practical choice, as it will collapse every time $m$ becomes a decimal. That's why all the codes below are hashtaged.

$$
g(y) = \frac{\nu e^{-\nu y}}{e^{\frac{\pi}{2}\nu} - e^{-\frac{\pi}{2}\nu}}
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = -\frac{1}{2}log[U \times (e^{\frac{-\pi}{2}\nu} - e^{\frac{\pi}{2}\nu}) + e^{\frac{\pi}{2}\nu}]
$$

```{r}
# # Generate Y
# n <- 10^5
# U <- runif(n)
# 
# Y <- lapply(1:length(nu), function(i){-0.5*log(U*(exp(-pi/2 * nu[i]) - exp(pi/2 * nu[i])) + exp(pi/2 * nu[i]))})
```

```{r}
# Generate the sample using AR
# U <- runif(n)
# 
# sample_envelope_2 <- lapply(1:length(nu), function(i){
#   Y[[i]][U < (cos(Y[[i]]))^{2*m[i]}]
# })
# 
# plot(density(sample_envelope_2[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
# X_envelope_2 <- lapply(1:length(nu), function(i) tan(sample_envelope_2[[i]]) * alpha[i] + lambda[i])
# 
# envelop_2_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_2[[i]])))
# envelop_2_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_2[[i]])))
```

```{r}
# # empirical rate of acceptance
# envelope_2_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_2[[i]])/10^5))
# ```
# ```{r}
# # update Q2_summary
# Q2_summary <- Q2_summary %>% 
#   mutate(envelop_2_empirical_mean, envelop_2_empirical_var, envelope_2_empirical_acceptance)
# 
# Q2_summary
```


#### Codes for Envelope function 3 (cosine)
$$
g(y) = \frac{1}{2}cos(y)
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = sin^{-1}(2U - 1)
$$

```{r include=FALSE}
# Generate Y
set.seed(1)
n <- 10^5
U <- runif(n)

Y <- asin(2*U-1)
sd(Y)/sqrt(n)
```

```{r include=FALSE}
# Control variate method
set.seed(1)
Y_prime <- 1/2 - U
c <- -12*cov(Y, Y_prime)
sample <- Y + c*Y_prime
sd(sample)/sqrt(n)
```

```{r include=FALSE}
# Generate the sample using AR
set.seed(1)
ptm <- proc.time()
U <- runif(n)

temp <- atan(-nu/(2*m-1))

sample_envelope_3 <- lapply(1:length(nu), function(i){
  temp <- temp[i]
  m <- m[i]
  nu <- nu[i]
  
  Y[U < (cos(Y)^{2*m-1}*exp(-nu*Y))/(cos(temp)^{2*m-1}*exp(-nu*temp))]
})
sim_envelope_3 <- (proc.time()- ptm)[3]

plot(density(sample_envelope_3[[1]]), type = 'l')
```

```{r include=FALSE}
# Produce X, and check mean and variance
X_envelope_3 <- lapply(1:length(nu), function(i) tan(sample_envelope_3[[i]]) * alpha[i] + lambda[i])

envelop_3_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_3[[i]])))
envelop_3_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_3[[i]])))
```

```{r include=FALSE}
# empirical rate of acceptance
envelope_3_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_3[[i]])/10^5))
```

```{r include=FALSE}
# update Q2_summary
Q2_summary <- Q2_summary %>% 
  mutate(envelop_3_empirical_mean, envelop_3_empirical_var, envelope_3_empirical_acceptance)
Q2_summary
```

#### Codes for Envelope function 4 (Cauchy)
$$
g(y) \propto (1+(\frac{y-\lambda}{\alpha})^{2})^{-1}
$$
With inverse transform algorithm, we get
$$
F^{-1}(U) = \alpha\times  tan(\pi U - \frac{\pi}{2}) + \lambda
$$

```{r include=FALSE}
# Generate Y
set.seed(1)
n <- 10^5
Y <- lapply(1:length(nu), function(i) alpha[i] * tan(pi*runif(n) - pi/2) + lambda[i])
```

```{r include=FALSE}
# Generate the sample using AR
set.seed(1)
ptm <- proc.time()
U <- runif(n)

temp_1 <- lapply(1:length(nu), function(i) (Y[[i]] - lambda[i])/alpha[i])
temp_2 <- nu/(2*(1 - m))

sample_envelope_4 <- lapply(1:length(nu), function(i){
  Y <- Y[[i]]
  temp_1 <- temp_1[[i]]
  temp_2 <- temp_2[i]
  nu <- nu[i]
  m <- m[i]
  
  Y[U < ((1 + temp_1^2)^{-m} * exp(-nu * atan(temp_1)))/((1 + temp_2^2)^{1-m} * exp(-nu*atan(temp_2)))]
})
sim_envelope_4 <- (proc.time()- ptm)[3]
```

```{r include=FALSE}
# Produce X, and check mean and variance
X_envelope_4 <- sample_envelope_4

envelop_4_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_4[[i]])))
envelop_4_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_4[[i]])))
```

```{r include=FALSE}
# empirical rate of acceptance
envelope_4_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_4[[i]])/10^5))
```

```{r include=FALSE}
# update summary_Q2
Q2_summary <- Q2_summary %>% 
  mutate(envelop_4_empirical_mean, envelop_4_empirical_var, envelope_4_empirical_acceptance)
```


#### Codes for Envelope function 5 (t distribution)
Our envelope function is \
$$
g(y) \propto (1+(\frac{y-\lambda}{\alpha})^{2})^{-m}
$$
We found it hard to find the exact expression of a constant that allows it to integrate to one (i.e. make it a PDF). \
We solved this problem by getting some inspirations from the PDF of t-distribution, i.e. $f(y) \propto (1+\frac{y^{2}}{k})^{-\frac{k+1}{2}}$. We transformed $g(y)$ to the following t-distribution form by setting $m = \frac{k+1}{2}$ and $z = \frac{\sqrt{k}(y - \lambda)}{\alpha}$: \
$$
g(y) \propto (1+\frac{z^{2}}{k})^{-\frac{k+1}{2}}
$$
As a result, we are able to simulate $Y$ indirectly by \
Step 1: Simulate a $Z$ from the built-in t-distribution function `rt` \
Step 2: Transform $Z$ to $Y$ by using $Y = Z \frac{\alpha}{\sqrt{k}} + \lambda$ \
```{r include=FALSE}
# Generate Y
set.seed(1)
n <- 10^5
k <- 2*m - 1

Y <- lapply(1:length(nu), function(i){
  rt(n, df = k[i]) * alpha[i]/sqrt(k[i]) + lambda[i]
})
```

```{r include=FALSE}
# Generate the sample using AR
set.seed(1)
ptm <- proc.time()
U <- runif(n)

sample_envelope_5 <- lapply(1:length(nu), function(i){
  Y <- Y[[i]]
  nu <- nu[i]
  lambda <- lambda[i]
  alpha <- alpha[i]
  
  Y[U < exp(-nu*atan((Y - lambda)/alpha) - abs(nu)*pi/2)]
})
sim_envelope_5 <- (proc.time()- ptm)[3]

plot(density(sample_envelope_5[[1]]), type = 'l')
```

```{r include=FALSE}
# Check mean and variance
X_envelope_5 <- sample_envelope_5

envelop_5_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_5[[i]])))
envelop_5_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_5[[i]])))
```

```{r include=FALSE}
# empirical rate of acceptance
envelope_5_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_5[[i]])/10^5))
```

```{r include=FALSE}
# update summary_Q2
Q2_summary <- Q2_summary %>% 
  mutate(envelop_5_empirical_mean, envelop_5_empirical_var, envelope_5_empirical_acceptance)
```



### Pick the best envelope function \
When writing the codes, we noticed that envelope function 2 is not a practical choice, as it will collapse every time $m$ becomes a decimal. Hence, we kicked it out from the further analysis of performances. \

To pick the best envelope function, we compare **a)** the empirical rate of acceptance and **b)** the time of running the acceptance-rejection algorithm. \

We summarise the empirical rate of acceptance of these envelope function in the boxplot below: \
```{r echo=FALSE}
Q2_summary_acceptance <- Q2_summary %>%
  select(m, nu, lambda, alpha, envelope_1 = envelope_1_empirical_acceptance, envelope_3 = envelope_3_empirical_acceptance, envelope_4 = envelope_4_empirical_acceptance, envelope_5 = envelope_5_empirical_acceptance)

df_long <- pivot_longer(Q2_summary_acceptance[, 5:ncol(Q2_summary_acceptance)], cols = everything(), names_to = "Envelope", values_to = "Acceptance")

# Plot
ggplot(df_long, aes(x = Envelope, y = Acceptance, fill = Envelope)) +
  geom_boxplot() +
  theme_minimal() + 
  theme(legend.position = "none") +
  labs(x = 'Envelope Function', y = 'Empirical Acceptance Rate', title = 'Compare the Performance of Different Envelope Functions')
```

We summarised the time consumed to run different algorithms in the following table: \
```{r echo=FALSE}
rbind(sim_envelope_1, sim_envelope_3, sim_envelope_4, sim_envelope_5)
```

We can see that generally, envelope 3 has significantly better acceptance rates than the others, while it has relatively good computational efficiency. Thus, pick envelope 3 as our envelope function. \

Below is extra details about the algorithm we used for envelope 3: \
With transformation $y = arctan(\frac{x - \lambda}{\alpha})$, we have \
$$
f(y) \propto (\cos(y))^{2m}e^{-\nu y} \\
g(y) = \frac{1}{2}\cos(y)
$$

Accept the sample from the envelope function when \
$$
U < \frac{(cos(y))^{2m - 1} e^{-\nu y}}{\left[cos\left( arctan(-\frac{\nu}{2m-1})\right)\right] e^{-\nu \ arctan(-\frac{\nu}{2m-1})}}
$$

The question also asks which values of parameters will give the most efficient simulation. We will answer this question here. \
By analysing the information from the updated `Q2_summary` dataframe with `dplyr` package, we are able to observe that the rate of acceptance is not affected by either $\lambda$ or $\alpha$. Thus, check the relationship between rate of acceptance and $m$, $\nu$, using `ggplot2`: \

```{r include=FALSE}
Q2_summary_acceptance_3 <- Q2_summary_acceptance %>%
  select(m, nu, lambda, alpha, envelope_3) %>%
  arrange(desc(envelope_3))

Q2_summary_acceptance_3
```

```{r include=FALSE}
Q2_summary_acceptance_temp <- Q2_summary_acceptance_3 %>% 
  group_by(envelope_3) %>%
  summarize(m = unique(m), nu = unique(nu)) %>%
  arrange(desc(envelope_3))

Q2_summary_acceptance_temp
```

```{r echo=FALSE}
ggplot(Q2_summary_acceptance_temp, aes(x = m, y = envelope_3, col = factor(nu))) + 
         geom_point(alpha = 0.5, size = 5) + 
         theme_minimal() +
         labs(y = 'Empirical Acceptance Rate', color = expression(nu), title = 'Empirical Acceptance Rates vs m')

ggplot(Q2_summary_acceptance_temp, aes(x = nu, y = envelope_3, col = factor(m))) + 
         geom_point(alpha = 0.5, size = 5) + 
         theme_minimal() + 
         labs(y = 'Empirical Acceptance Rate', x = expression(nu), color = 'm', title = expression('Empirical Acceptance Rates vs '*nu))
```

From the plot, we can see that the acceptance rate is highest when $m$ is small, **AND** $\nu$ is close to zero.

### b)
We have already calculated the theoretical mean and variances when we defined `Q2_summary` early on. Now, we compare them with the sample mean and variances using the `dplyr` package, and the differences between them are plotted. 
From the dataframe, we noticed that $\lambda$ has no effect on either the deviation from theoretical mean or the deviation from theoretical variances. Use plots to investigate the effects of other parameters:

#### Analysis on mean
```{r include=FALSE}
Q2_summary_empirical_mean <- Q2_summary %>%
  mutate(diff = envelop_3_empirical_mean - theoretical_mean) %>%
  select(m, nu, lambda, alpha, diff) %>%
  arrange(diff)

Q2_summary_empirical_mean
mean(Q2_summary_empirical_mean$diff) # the mean difference is very small
```

```{r include=FALSE}
Q2_summary_empirical_mean_temp <- Q2_summary_empirical_mean %>% 
  group_by(diff) %>%
  summarize(m = unique(m), nu = unique(nu), alpha = unique(alpha)) %>%
  arrange(diff)
```

#### Analysis on variance
```{r include=FALSE}
Q2_summary_empirical_var <- Q2_summary %>%
  mutate(diff = envelop_3_empirical_var - theoretical_var) %>%
  select(m, nu, lambda, alpha, diff) %>%
  arrange(diff)

mean(Q2_summary_empirical_var$diff) # the difference to theoretical variance is large
sum(Q2_summary_empirical_var$diff >=0) # there is no empirical variance that is higher than the theoretical variance.
```

```{r include=FALSE}
Q2_summary_empirical_var_temp <- Q2_summary_empirical_var %>% 
  group_by(diff) %>%
  summarize(m = unique(m), nu = unique(nu), alpha = unique(alpha)) %>%
  arrange(diff)
```

#### Plots
```{r echo=FALSE}
# Mean plots
p_mean_m <- ggplot(Q2_summary_empirical_mean_temp, aes(m, diff)) +
  geom_point() + 
  theme_minimal() +
  labs(y = 'Sample Mean - Theoretical Mean', title = 'Difference in mean vs m') +
  theme(axis.title.y = element_text(size = 7), title = element_text(size = 8))
p_mean_nu <- ggplot(Q2_summary_empirical_mean_temp, aes(nu, diff)) + 
  geom_point() + 
  theme_minimal() +
  labs(y = 'Sample Mean - Theoretical Mean', x = expression(nu), title = expression('Difference in mean vs '*nu)) +
  theme(axis.title.y = element_text(size = 7), title = element_text(size = 8))
p_mean_alpha <- ggplot(Q2_summary_empirical_mean_temp, aes(alpha, diff)) + 
  geom_point() + 
  theme_minimal() +
  labs(y = 'Sample Mean - Theoretical Mean', x = expression(alpha), title = expression('Difference in mean vs '*alpha)) +
  theme(axis.title.y = element_text(size = 7), title = element_text(size = 8))

# Variance plots
p_var_m <- ggplot(Q2_summary_empirical_var_temp, aes(m, diff)) + 
  geom_point() + 
  theme_minimal() +
  labs(y = 'Sample Variance - Theoretical Variance', title = 'Difference in variance vs m') +
  theme(axis.title.y = element_text(size = 7), title = element_text(size = 8))
p_var_nu <- ggplot(Q2_summary_empirical_var_temp, aes(nu, diff)) + 
  geom_point() + 
  theme_minimal() +
  labs(y = 'Sample Variance - Theoretical Variance', x = expression(nu), title = expression('Difference in variance vs '*nu)) +
  theme(axis.title.y = element_text(size = 7), title = element_text(size = 8))
p_var_alpha <- ggplot(Q2_summary_empirical_var_temp, aes(alpha, diff)) + 
  geom_point() + 
  theme_minimal() +
  labs(y = 'Sample Variance - Theoretical Variance', x = expression(alpha), title = expression('Difference in variance vs '*alpha)) +
  theme(axis.title.y = element_text(size = 7), title = element_text(size = 8))

grid.arrange(p_mean_m, p_mean_nu, p_mean_alpha, p_var_m, p_var_nu, p_var_alpha, ncol = 3)
```

From the plots and the other analysis from the dataframe, we draw the following conclusions: \
**1.** The differences between the sample and theoretical mean/variance is the smallest when \
- $\m$ is large; \
- $\nu$ is close to zero; \
- $\alpha$ is small. \
**2.** The mean of all differences in mean is small. \
**3.** The mean of all differences in variance is large, and there is no empirical variance that is higher than the theoretical variance. \\

### c)
#### Empirical quantiles
```{r}
length(X_envelope_3)
quantile_percents <- c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)

matrix <- matrix(numeric(length = length(nu)*length(quantile_percents)), nrow = length(nu))

for (i in 1:length(nu)){
  matrix[i,] <- quantile(X_envelope_3[[i]], prob = quantile_percents)
}

dim(matrix)
```

```{r}
Q2_empirical_quantiles <- as.data.frame(matrix) %>% mutate(index = 1:length(nu))
colnames(Q2_empirical_quantiles) <- c(paste(as.character(quantile_percents*100), '%', sep = ''), 'index')

Q2_summary_empirical_quantiles <- Q2_summary %>% 
  mutate(index = 1:length(nu)) %>%
  select(m, nu, lambda, alpha, index, theoretical_mean, theoretical_var) %>%
  left_join(Q2_empirical_quantiles, by = join_by(index)) %>%
  select(-index)

Q2_summary_empirical_quantiles
```


### CLT quantiles
```{r}

```

## d)
### Redefine the parameters
```{r}
# Set m to be a large value. Also set arbitrary parameters sigma and lambda.
m <- 100
sigma <- 2
lambda <- 1

# Set the other parameters according to requirements
nu <- 0
alpha <- sigma*sqrt(2*m - 3)
alpha
```

### Generate sample from envelope 3 (cosine)
```{r}
# Generate Y
n <- 10^5
U <- runif(n)

Y <- asin(2*U-1)
```

```{r}
# Generate the sample using AR
U <- runif(n)

temp <- atan(-nu/(2*m-1))

Q2d_sample <- Y[U < (cos(Y)^{2*m-1}*exp(-nu*Y))/(cos(temp)^{2*m-1}*exp(-nu*temp))]
```

```{r}
# Produce X, and check mean, variance, acceptance
Q2d_X <- tan(Q2d_sample) * alpha + lambda

envelop_3_empirical_mean <- mean(Q2d_X)
envelop_3_empirical_var <- var(Q2d_X)
envelope_3_empirical_acceptance <- length(Q2d_X)/n
envelope_3_empirical_acceptance
```

### Use graphs to compare the sample with N(lambda, sigma)
```{r}
# PDF
x <- seq(min(Q2d_X), max(Q2d_X), 0.01)
plot(x, dnorm(x, mean = lambda, sd = sigma), type = 'l', main = 'Compare the PDF of the generated X and N(lambda, sigma^2)', ylab = 'density', col = 'red')
lines(density(Q2d_X), col = 'black')
legend('topright', legend = c('generated X', 'N(lambda, sigma^2)'), col = c('black', 'red'), lty = 1)
```

```{r}
# Q-Q plot
qqplot(rnorm(length(Q2d_X), mean = lambda, sd = sigma), Q2d_X, main = 'Q-Q plot of the generated X vs N(lambda, sigma^2)')
abline(a = 0, b = 1, col = 'red', lwd = 2)
legend('bottomright', legend = '45 degree reference line', col = 'red', lty = 1, lwd = 2)
```