---
title: "Q2"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 2
## a)
### Generate various values of m, nu, lambda and alpha
```{r}
m_list <- seq(1.6, 10, length.out = 10)
nu_list <- seq(-5, 5, length.out = 10)
lambda_list <- seq(-5, 5, length.out = 10)
alpha_list <- seq(0.1, 10, length.out = 10)

groups <- expand.grid(m_list, nu_list, lambda_list, alpha_list)

m <- groups$Var1
nu <- groups$Var2
lambda <- groups$Var3
alpha <- groups$Var4
```


```{r}
# m <- 2
# nu <- -1
# lambda <- -1
# alpha <- 1
```



A temporary check function:
```{r}
theoretical_mean <- lambda - alpha*nu/(2*(m - 1))
theoretical_var <- alpha^2/(2*m - 3) * (1 + nu^2/(4*(m - 1)^2))
```


```{r}
Q2_summary <- data.frame(m, nu, lambda, alpha, theoretical_mean, theoretical_var)
```



### With transformation
$$
u = tan^{-1}(\frac{x - \lambda}{\alpha}), u \in [-\frac{\pi}{2}, \frac{\pi}{2}]
$$
We get
$$
f(u) \propto (cos(u))^{2m}e^{-\nu u}
$$




#### Envelope function 1 (uniform distribution)
$$
g(y) = \frac{1}{\pi}, Y \sim Unif[-\frac{\pi}{2}, \frac{\pi}{2}]
$$

```{r}
# Generate Y
n <- 10^5
Y <- runif(n) * pi - pi/2
```

```{r}
# Generate the sample using AR
ptm <- proc.time()
U <- runif(n)

sample_envelope_1 <- lapply(1:length(nu), function(i) {
  m <- m[i]
  nu <- nu[i]
  
  temp <- atan(-nu / (2 * m))
  accepted_Y <- Y[U < 
    (cos(Y))^(2 * m) * exp(-nu * Y) / (cos(temp)^(2 * m) * exp(-nu * temp))]
})
sim_envelope_1 <- (proc.time()- ptm)[3]

plot(density(sample_envelope_1[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_1 <- lapply(1:length(nu), function(i){tan(sample_envelope_1[[i]]) * alpha[i] + lambda[i]})
# X_envelope_1[[1]][1:10]
# sample_envelope_1[[1]][1:10]

envelop_1_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_1[[i]])))
envelop_1_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_1[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_1_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_1[[i]])/10^5))
```

```{r}
# add the results to the summary
library(tidyverse)
Q2_summary <- Q2_summary %>% 
  mutate(envelop_1_empirical_mean, envelop_1_empirical_var, envelope_1_empirical_acceptance)

Q2_summary
```

#### Envelope function 2 (exponential)
$$
g(y) = \frac{\nu e^{-\nu y}}{e^{\frac{\pi}{2}\nu} - e^{-\frac{\pi}{2}\nu}}
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = -\frac{1}{2}log[U \times (e^{\frac{-\pi}{2}\nu} - e^{\frac{\pi}{2}\nu}) + e^{\frac{\pi}{2}\nu}]
$$

```{r}
# # Generate Y
# n <- 10^5
# U <- runif(n)
# 
# Y <- lapply(1:length(nu), function(i){-0.5*log(U*(exp(-pi/2 * nu[i]) - exp(pi/2 * nu[i])) + exp(pi/2 * nu[i]))})
```

```{r}
# Generate the sample using AR
# U <- runif(n)
# 
# sample_envelope_2 <- lapply(1:length(nu), function(i){
#   Y[[i]][U < (cos(Y[[i]]))^{2*m[i]}]
# })
# 
# plot(density(sample_envelope_2[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
# X_envelope_2 <- lapply(1:length(nu), function(i) tan(sample_envelope_2[[i]]) * alpha[i] + lambda[i])
# 
# envelop_2_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_2[[i]])))
# envelop_2_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_2[[i]])))
```

```{r}
# # empirical rate of acceptance
# envelope_2_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_2[[i]])/10^5))
# ```
# ```{r}
# # update Q2_summary
# Q2_summary <- Q2_summary %>% 
#   mutate(envelop_2_empirical_mean, envelop_2_empirical_var, envelope_2_empirical_acceptance)
# 
# Q2_summary
```


#### Envelope function 3 (cosine)
$$
g(y) = \frac{1}{2}cos(y)
$$

To generate Y, use inverse transform algorithm:
$$
F^{-1}(U) = sin^{-1}(2U - 1)
$$

```{r}
# Generate Y
n <- 10^5
U <- runif(n)

Y <- asin(2*U-1)
```

```{r}
# Generate the sample using AR
ptm <- proc.time()
U <- runif(n)

temp <- atan(-nu/(2*m-1))

sample_envelope_3 <- lapply(1:length(nu), function(i){
  temp <- temp[i]
  m <- m[i]
  nu <- nu[i]
  
  Y[U < (cos(Y)^{2*m-1}*exp(-nu*Y))/(cos(temp)^{2*m-1}*exp(-nu*temp))]
})
sim_envelope_3 <- (proc.time()- ptm)[3]

plot(density(sample_envelope_3[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_3 <- lapply(1:length(nu), function(i) tan(sample_envelope_3[[i]]) * alpha[i] + lambda[i])

envelop_3_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_3[[i]])))
envelop_3_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_3[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_3_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_3[[i]])/10^5))
```

```{r}
# update Q2_summary
Q2_summary <- Q2_summary %>% 
  mutate(envelop_3_empirical_mean, envelop_3_empirical_var, envelope_3_empirical_acceptance)
```

### Without transformation
#### Envelope function 4 (Cauchy)

```{r}
# Generate Y
n <- 10^5
Y <- lapply(1:length(nu), function(i) alpha[i] * tan(pi*runif(n) - pi/2) + lambda[i])
```

```{r}
# Generate the sample using AR
ptm <- proc.time()
U <- runif(n)

temp_1 <- lapply(1:length(nu), function(i) (Y[[i]] - lambda[i])/alpha[i])
temp_2 <- nu/(2*(1 - m))

sample_envelope_4 <- lapply(1:length(nu), function(i){
  Y <- Y[[i]]
  temp_1 <- temp_1[[i]]
  temp_2 <- temp_2[i]
  nu <- nu[i]
  m <- m[i]
  
  Y[U < ((1 + temp_1^2)^{-m} * exp(-nu * atan(temp_1)))/((1 + temp_2^2)^{1-m} * exp(-nu*atan(temp_2)))]
})
sim_envelope_4 <- (proc.time()- ptm)[3]


plot(density(sample_envelope_4[[1]]), type = 'l')
```

```{r}
# Produce X, and check mean and variance
X_envelope_4 <- sample_envelope_4

envelop_4_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_4[[i]])))
envelop_4_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_4[[i]])))
```

```{r}
# empirical rate of acceptance
envelope_4_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_4[[i]])/10^5))
```

```{r}
# update summary_Q2
Q2_summary <- Q2_summary %>% 
  mutate(envelop_4_empirical_mean, envelop_4_empirical_var, envelope_4_empirical_acceptance)

colnames(Q2_summary)
```


#### Envelope function 5 (t distribution)
```{r}
# Generate Y
n <- 10^5
k <- 2*m - 1

Y <- lapply(1:length(nu), function(i){
  rt(n, df = k[i]) * alpha[i]/sqrt(k[i]) + lambda[i]
})
```

```{r}
# Generate the sample using AR
ptm <- proc.time()
U <- runif(n)

sample_envelope_5 <- lapply(1:length(nu), function(i){
  Y <- Y[[i]]
  nu <- nu[i]
  lambda <- lambda[i]
  alpha <- alpha[i]
  
  Y[U < exp(-nu*atan((Y - lambda)/alpha) - abs(nu)*pi/2)]
})
sim_envelope_5 <- (proc.time()- ptm)[3]

plot(density(sample_envelope_5[[1]]), type = 'l')
```

```{r}
# Check mean and variance
X_envelope_5 <- sample_envelope_5

envelop_5_empirical_mean <- unlist(lapply(1:length(nu), function(i) mean(X_envelope_5[[i]])))
envelop_5_empirical_var <- unlist(lapply(1:length(nu), function(i) var(X_envelope_5[[i]])))
```


```{r}
# empirical rate of acceptance
envelope_5_empirical_acceptance <- unlist(lapply(1:length(nu), function(i) length(sample_envelope_5[[i]])/10^5))
```

```{r}
# update summary_Q2
Q2_summary <- Q2_summary %>% 
  mutate(envelop_5_empirical_mean, envelop_5_empirical_var, envelope_5_empirical_acceptance)
```



### Export summary_Q2
```{r}
#write.csv(Q2_summary, 'Q2_summary_10_param.csv', row.names = F)
```

### Check the time consumed
```{r}
#rbind(sim_envelope_1, sim_envelope_3, sim_envelope_4, sim_envelope_5)
sim_envelope_3
```
### Analysis on acceptance
```{r}
Q2_summary_acceptance <- Q2_summary %>%
  select(m, nu, lambda, alpha, envelope_3 = envelope_3_empirical_acceptance) %>%
  arrange(desc(envelope_3))

length(unique(Q2_summary_acceptance$envelope_3))
Q2_summary_acceptance
```

It is clear that the rate of acceptance is not affected by either $\lambda$ or $\alpha$. Thus, check the relationship between rate of acceptance and $m$, $\nu$:

```{r}
Q2_summary_acceptance_temp <- Q2_summary_acceptance %>% 
  group_by(envelope_3) %>%
  summarize(m = unique(m), nu = unique(nu)) %>%
  arrange(desc(envelope_3))

Q2_summary_acceptance_temp
```

```{r}
# Try ggplot2
library(ggplot2)
ggplot(Q2_summary_acceptance_temp, aes(x = m, y = envelope_3, col = factor(nu))) + 
         geom_point(alpha = 0.5, size = 5) + 
         theme_minimal()

ggplot(Q2_summary_acceptance_temp, aes(x = nu, y = envelope_3, col = factor(m))) + 
         geom_point(alpha = 0.5, size = 5) + 
         theme_minimal()
```

Hence, the acceptance rate is highest when $m$ is small, **AND** $\nu$ is close to zero.

## b)
### Analysis on mean
```{r}
Q2_summary_empirical_mean <- Q2_summary %>%
  mutate(diff = envelop_3_empirical_mean - theoretical_mean) %>%
  select(m, nu, lambda, alpha, diff) %>%
  arrange(diff)

Q2_summary_empirical_mean
mean(Q2_summary_empirical_mean$diff) # the mean difference is very small
```

It is obvious that $\lambda$ has no effect on the deviation from theoretical mean. Use plots to investigate the effects of other parameters:
```{r}
Q2_summary_empirical_mean_temp <- Q2_summary_empirical_mean %>% 
  group_by(diff) %>%
  summarize(m = unique(m), nu = unique(nu), alpha = unique(alpha)) %>%
  arrange(diff)
```

```{r}
# Plots
ggplot(Q2_summary_empirical_mean_temp, aes(m, diff, col = factor(nu))) + geom_point() + theme_minimal()
ggplot(Q2_summary_empirical_mean_temp, aes(nu, diff, col = factor(m))) + geom_point() + theme_minimal()
ggplot(Q2_summary_empirical_mean_temp, aes(alpha, diff, col = factor(nu))) + geom_point() + theme_minimal()
```

Hence, the difference to the theoretical mean is the smallest when \\
- $\m$ is large; \\
- $\nu$ is close to zero; \\
- $\alpha$ is small.\\

### Analysis on variance
```{r}
Q2_summary_empirical_var <- Q2_summary %>%
  mutate(diff = envelop_3_empirical_var - theoretical_var) %>%
  select(m, nu, lambda, alpha, diff) %>%
  arrange(diff)

Q2_summary_empirical_var
mean(Q2_summary_empirical_var$diff) # the difference to theoretical variance is large
sum(Q2_summary_empirical_var$diff >=0) # there is no empirical variance that is higher than the theoretical variance.
```

It is obvious that $\lambda$ has no effect on the deviation from theoretical variance. Use plots to investigate the effects of other parameters:
```{r}
Q2_summary_empirical_var_temp <- Q2_summary_empirical_var %>% 
  group_by(diff) %>%
  summarize(m = unique(m), nu = unique(nu), alpha = unique(alpha)) %>%
  arrange(diff)
```

```{r}
# Plots
ggplot(Q2_summary_empirical_var_temp, aes(m, diff, col = factor(nu))) + geom_point() + theme_minimal()
ggplot(Q2_summary_empirical_var_temp, aes(nu, diff, col = factor(m))) + geom_point() + theme_minimal()
ggplot(Q2_summary_empirical_var_temp, aes(alpha, diff, col = factor(nu))) + geom_point() + theme_minimal()
```

Hence, the difference to the theoretical variance is the smallest when \\
- $\m$ is large; \\
- $\nu$ is close to zero; \\
- $\alpha$ is small.\\
Is this a good thing or not?

## c)
### Empirical quantiles
```{r}
length(X_envelope_3)
quantile_percents <- c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)

matrix <- matrix(numeric(length = length(nu)*length(quantile_percents)), nrow = length(nu))

for (i in 1:length(nu)){
  matrix[i,] <- quantile(X_envelope_3[[i]], prob = quantile_percents)
  print(i)
}
```

```{r}
Q2_empirical_quantiles <- as.data.frame(matrix) %>% mutate(index = 1:length(nu))
colnames(Q2_empirical_quantiles) <- c(paste(as.character(quantile_percents*100), '%', sep = ''), 'index')
Q2_empirical_quantiles
```

```{r}
Q2_summary_empirical_quantiles <- Q2_summary %>% 
  mutate(index = 1:length(nu)) %>%
  select(m, nu, lambda, alpha, index, theoretical_mean, theoretical_var) %>%
  left_join(Q2_empirical_quantiles, by = join_by(index)) %>%
  select(-index)

Q2_summary_empirical_quantiles
```

### CLT quantiles
```{r}

```

## d)
### Redefine the parameters
```{r}
# Set m to be a large value. Also set arbitrary parameters sigma and lambda.
m <- 10000
sigma <- 2
lambda <- 1

# Set the other parameters according to requirements
nu <- 0
alpha <- sigma*sqrt(2*m - 3)
alpha
```

### Generate sample from envelope 3 (cosine)
```{r}
# Generate Y
n <- 10^5
U <- runif(n)

Y <- asin(2*U-1)
```

```{r}
# Generate the sample using AR
U <- runif(n)

temp <- atan(-nu/(2*m-1))

Q2d_sample <- Y[U < (cos(Y)^{2*m-1}*exp(-nu*Y))/(cos(temp)^{2*m-1}*exp(-nu*temp))]
```

```{r}
# Produce X, and check mean, variance, acceptance
Q2d_X <- tan(Q2d_sample) * alpha + lambda

envelop_3_empirical_mean <- mean(Q2d_X)
envelop_3_empirical_var <- var(Q2d_X)
envelope_3_empirical_acceptance <- length(Q2d_X)/n
envelope_3_empirical_acceptance
```

### Use graphs to compare the sample with N(lambda, sigma)
```{r}
# PDF
x <- seq(min(Q2d_X), max(Q2d_X), 0.01)
plot(x, dnorm(x, mean = lambda, sd = sigma), type = 'l', main = 'Compare the PDF of the generated X and N(lambda, sigma^2)', ylab = 'density', col = 'red')
lines(density(Q2d_X), col = 'black')
legend('topright', legend = c('generated X', 'N(lambda, sigma^2)'), col = c('black', 'red'), lty = 1)
```

```{r}
# Q-Q plot
qqplot(rnorm(length(Q2d_X), mean = lambda, sd = sigma), Q2d_X, main = 'Q-Q plot of the generated X vs N(lambda, sigma^2)')
abline(a = 0, b = 1, col = 'red', lwd = 2)
legend('bottomright', legend = '45 degree reference line', col = 'red', lty = 1, lwd = 2)
```