---
title: "ST303 group project"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2025-04-04"
---

## Question 3

### (a)

A Bessel process $R_t$ of order $\delta \geq 1$ has the SDE:

$$
dR_t = \frac{\delta - 1}{2R_t} \, dt + dW_t, \quad R_0 = 1
$$

Consider the SDE for the squared Bessel process $Y_t = R_t^2$, by Itô's formula:

$$
dY_t = 2R_t \, dR_t + \frac{1}{2} \cdot 2 \, d[R]_t, \quad \text{where } d[R]_t = dt
$$

$$
= 2R_t \left( \frac{\delta - 1}{2R_t} \, dt + dW_t \right) + dt
$$

$$
= \delta \, dt + 2 \sqrt{Y_t} \, dW_t,  \quad dW_t \sim \sqrt{dt} \cdot \mathcal{N}(0, 1)
$$

$$
Y_t = Y_0 + \int_0^t \delta \, ds + \int_0^t 2\sqrt{Y_s} \, dW_s, \quad Y_0 = R_0^2 = 1
$$

Check:

$$
\mathbb{E}(Y_t) = Y_0 + \delta t, \quad \text{since } \mathbb{E} \left( \int_0^t 2\sqrt{Y_s} \, dW_s \right) = 0
$$


```{r}
set.seed(1)
T = 10
N = 1000
dt = T/N
Rt_sim <- function(delta, N, dt){
  Y <- numeric(N+1)
  Y[1] <- 1
  for (i in 1:N){
    dW <- sqrt(dt)*rnorm(1)
    Y[i+1] <- max(Y[i] + delta*dt+ 2*sqrt(Y[i])*dW, 0)
  } 
  return(sqrt(Y))
}
t <- seq(0, T, by = dt)
delta <- c(1,2,3,4)
Rt1 <- Rt_sim(delta[1], N=1000, dt=10/1000)
Rt2 <- Rt_sim(delta[2], N=1000, dt=10/1000)
Rt3 <- Rt_sim(delta[3], N=1000, dt=10/1000)
Rt4 <- Rt_sim(delta[4], N=1000, dt=10/1000)
ymin <- min(c(Rt1, Rt2, Rt3, Rt4))
ymax <- max(c(Rt1, Rt2, Rt3, Rt4))
plot(t, Rt1, "l", col = "yellow", ylim = c(ymin, ymax), main = "Sample Paths of Bessel Process", xlab = "Time t", ylab = expression(R[t]))
lines(t, Rt2, "l", col="red")
lines(t, Rt3, "l", col="blue")
lines(t, Rt4, "l", col="green")
legend("topleft",
       legend = c(expression(delta *' = 1'), expression(delta *' = 2'), expression(delta *' = 3'), expression(delta *' = 4')),
       col = c("yellow", "red", "blue", "green"),
       lty = 1)  
```

### (b)

Generate more paths over $t \in [0, 10]$ with $\delta=1$:

```{r}
set.seed(1)
M <- 100  # Number of sample paths
t <- seq(0, T, by = dt)
# choose delta = 1, generate more paths
Rt_matrix <- sapply(1:M, function(m) Rt_sim(delta[1], N, dt)) # row: dt, col: Rt
matplot(t, Rt_matrix, type = "l", lty = 1, col = rainbow(M),
        main = "Sample Paths of Bessel Process (δ = 1)",
        xlab = "Time t", ylab = expression(R[t]))
```

Simulate maximum for each path and calculate the mean of this maximum:

```{r}
max_Rt <- apply(Rt_matrix, 2, max) # calculate maximum for each column of Rt_matrix
hist(max_Rt, breaks = 20, prob = T, 
     main = expression("Histogram of Max " * R[t] * " over [0,10] (δ = 1)"), 
     xlab = expression("Max " * R[t]))
mean(max_Rt)
legend("topright", legend = "mean = 3.812538")
```

### (c)

Consider a Bermudan call option with the following characteristics:

- Underlying stock price: $R_t$, a Bessel process with order $\delta = 1.2$
- Strike price: $K = 2$
- Maturity: $T = 10$
- Exercisable at discrete times: $t_k = k$, for $k = 1, 2, \dots, 10$
- Exercise time: $\tau$ is a stopping time chosen from $\{t_1, t_2, \dots, t_{10}\}$
- Payoff at exercise: $(R_\tau - K)^+ = \max(R_\tau - K, 0)$

First, generate sample paths for stock price $R_t$ with $\delta=1.2$.

```{r}
set.seed(1)
delta = 1.2
K = 2
T = 10
M <- 100  # Number of sample paths
t <- seq(0, T, by = dt)
Rt_matrix <- sapply(1:M, function(m) Rt_sim(delta, N, dt)) # rows: t, cols: mth path of Rt
matplot(t, Rt_matrix, type = "l", lty = 1, col = rainbow(M),
        main = "Sample Paths of Bessel Process (δ = 1.2)",
        xlab = "Time t", ylab = expression(R[t]))
```

Second, construct payoff matrix for exercisable dates.
The option will be exercised when the highest payoff is reached. Assign this values to the exercise time $\tau$.

```{r}
tk_index <- (1:10)/(10/1000) + 1
Rtk <- Rt_matrix[tk_index, ]  # rows: tk, cols: mth path of Rt
tk_payoff <- pmax(Rtk - K, 0)  # (Rt - K)+
exercise_time <- apply(tk_payoff, 2, function(payoffs) {
  idx <- which.max(payoffs)  # highest positive payoff
  if (is.na(idx)) return(NA)   # never exercised
  return(idx)  # tk index 1:10
})
# Convert index to actual time τ
tau <- ifelse(is.na(exercise_time), NA, exercise_time)  # τ in 1:10
hist(tau, prob = T, breaks = 20)
lines(density(tau[!is.na(tau)]))
summary(tau) # median = 8, mean = 6.89
```

In conclusion, the result(median=8,mean=6.89) shows that there are no incentives to exercise early. Since there are no discounting factors, most options are exercised close to maturity.

### (d)

Assuming no discounting factor, i.e. $r=0$ thus $e^{-rt}=1$, the pricing formula of the Bermuda call option is $E(max((R_{t_{i}}-K)^+, E(V_{i+1}|Fi)))$, where $V_{i+1}$ is the future value. Hence this is equivalent to find the maximum value of the payoffs all along the exercisable times, i.e. $max((R_{t_1}-K)^+,..., (R_{t_{10}}-K)^+)$

```{r}
Bermuda_payoffs <- sapply(1:M, function(m) {
  payoffs <- tk_payoff[, m]
  holding_values <- numeric(10)
  holding_values[10] <- payoffs[10]
  for (t in 9:1) {
    if (payoffs[t] > 0) {
      immediate_exercise <- payoffs[t]
      future_payoff <- holding_values[t + 1]
      holding_values[t] <- max(immediate_exercise, future_payoff)
    } else {
      holding_values[t] <- holding_values[t + 1]
    }
  }
  return(holding_values[1])
})
Bermuda_price <- mean(Bermuda_payoffs)
cat("Bermuda call price =", Bermuda_price, "\n")
```

The European call option price $= E((R_{10} - K)^+)$.

```{r}
# European call option price = E((R10-K)+)
European_payoff <- tk_payoff[10,]
European_price <- mean(European_payoff)
cat("European call price =", European_price, "\n")
```

Calculate how much more expensive Bermuda call option is than an European
call option:

```{r}
diff_price = Bermuda_price - European_price
cat("Extra cost of Bermuda over European =", diff_price, "\n")
```

### (e)

```{r}
# 1. No proper modeling of expected future values. (we just take average: mean())
# 2. Discounting factor = 0 here, there is little incentive to exercise early. (if r>0, price = e^-rt*E(Vi+1))
# 3. Y[i+1] <- max(Y[i] + delta*dt+ 2*sqrt(Y[i])*dW, 0)
# 4. Variance reduction: AV, CV, conditioning, stratify sampling, importance sampling
```

#### (1)
We used a single “average of future payoffs” or just “the next time step’s payoff,” that might be a coarse approximation.
The more accurate method in continuous models is to do a regression across all paths’ next‐step (or multiple‐step) payoffs on the underlying Rt.

#### (2)
In real-world financial markets, the valuation of options typically incorporates a **positive risk-free interest rate** \( r > 0 \), which introduces a discounting factor \( e^{-rt} \) into the pricing model. This affects the **expected present value** of future payoffs, and has direct consequences for the early exercise behavior in Bermuda options.

Under discounting, the value of the Bermuda option at any exercisable time \( t_i \) is given by:

\[
V_i = \max\left( (R_{t_i} - K)^+, \ e^{-r \Delta t} \cdot \mathbb{E}[V_{i+1} \mid \mathcal{F}_{t_i}] \right)
\]

Here, \( \Delta t \) is the time between exercise opportunities, and \( \mathcal{F}_{t_i} \) denotes the information available at time \( t_i \). The exponential discounting term \( e^{-r \Delta t} \) reduces the value of waiting, thereby making **early exercise more attractive** in some scenarios.

This dynamic alters the option holder’s optimal stopping strategy. Instead of always deferring exercise in hopes of higher future payoffs, the holder may choose to exercise earlier to avoid the loss in present value due to discounting. Consequently, the **Bermuda option becomes strictly more valuable than the European option**, which only allows exercise at maturity.

In summary, introducing a positive discount rate \( r > 0 \) changes the valuation structure of the Bermuda option by lowering the continuation value through discounting. This increases the importance of early exercise opportunities and typically results in a higher option price compared to the no-discounting scenario.

#### (3) 
$$
    R_t \to 0 \quad \Rightarrow \quad Y_t = R_t^2 \to 0
$$ 
To avoid taking the square root of negative values when simulating $R_t = \sqrt{Y_t}$, we applied a non-negativity constraint in the simulation of $Y_t$:

$$
Y_{i+1} = \max\left( Y_i + \delta \cdot \Delta t + 2 \sqrt{Y_i} \cdot \Delta W_i,\ 0 \right)
$$

This ensures that $Y_t \geq 0$, so that $R_t = \sqrt{Y_t}$ is always real-valued.


#### (4) Variance Reduction:

Control Variate Method:

The price of the Bermuda call option is equal to $max((R_{t_1}-K)^+,..., (R_{t_{10}}-K)^+)$, which is equivalent to $max(max(R_t)-K,\ 0)$. max(R_t)

Let $X = max(max(R_t)-K,\ 0), \ for\ t=1,2,...,10$. Set $\tilde{X} = max(R_t)-K$, then $Y = \tilde{X} - E(\tilde{X})$. Check: $E(Y) = 0$.

```{r}
max_Rtk <- apply(Rtk, 2, max)
X_tilde = max_Rtk - K
X = pmax(X_tilde, 0)
Y = X_tilde - mean(X_tilde)
mean(Y) # check E(Y)=0
```

Calculate $c=-\frac{Cov(X,Y)}{Var(Y)}$:

```{r}
c <- -cov(X,Y)/var(Y)
c
```

Then $\hat{\theta}=X + cY$, calculate $Var(\hat{\theta})$ and compare it to $Var(X)$:
```{r}
theta_hat = X + c*Y
cat(var(X), var(theta_hat), "\n",  
    "Improved by",(var(X)-var(theta_hat))/var(X) * 100, "%.")
```

Other attempts for variance reduction:

1. Antithetic Variates Method: unavailable since $R_t$ is not a monotone function.
2. Stratify Sampling: unavailable since inverse transformed estimator is needed.
3. Conditioning Sampling: unavailable since $R_t$ is the only variable.
4. Importance Sampling: unavailable since the distribution of $R_t$ is not given.




