---
title: "ST303 group project"
output: html_document
date: "2025-04-04"
---

## Question 3

### (a)
```{r}
set.seed(1)
T = 10
N = 1000
dt = T/N
Rt_sim <- function(delta, N, dt){
  Y <- numeric(N+1)
  Y[1] <- 1
  for (i in 1:N){
    dW <- sqrt(dt)*rnorm(1)
    Y[i+1] <- max(Y[i] + delta*dt+ 2*sqrt(Y[i])*dW, 0)
  } 
  return(sqrt(Y))
}
t <- seq(0, T, by = dt)
delta <- c(1,2,3,4)
Rt1 <- Rt_sim(delta[1], N=1000, dt=10/1000)
Rt2 <- Rt_sim(delta[2], N=1000, dt=10/1000)
Rt3 <- Rt_sim(delta[3], N=1000, dt=10/1000)
Rt4 <- Rt_sim(delta[4], N=1000, dt=10/1000)
ymin <- min(c(Rt1, Rt2, Rt3, Rt4))
ymax <- max(c(Rt1, Rt2, Rt3, Rt4))
plot(t, Rt1, "l", col = "yellow", ylim = c(ymin, ymax), main = "Sample Paths of Bessel Process", xlab = "Time t", ylab = expression(R[t]))
lines(t, Rt2, "l", col="red")
lines(t, Rt3, "l", col="blue")
lines(t, Rt4, "l", col="green")
legend("topleft",
       legend = c(expression(delta *' = 1'), expression(delta *' = 2'), expression(delta *' = 3'), expression(delta *' = 4')),
       col = c("yellow", "red", "blue", "green"),
       lty = 1)  
```

### (b)
```{r}
set.seed(1)

M <- 100  # Number of sample paths
t <- seq(0, T, by = dt)

# choose delta = 1, generate more paths
Rt_matrix <- sapply(1:M, function(m) Rt_sim(delta[1], N, dt)) # row: dt, col: Rt

matplot(t, Rt_matrix, type = "l", lty = 1, col = rainbow(M),
        main = "Sample Paths of Bessel Process (δ = 1)",
        xlab = "Time t", ylab = expression(R[t]))

max_Rt <- apply(Rt_matrix, 2, max) # calculate maximum for each column of Rt_matrix

hist(max_Rt, breaks = 20, prob = T, 
     main = expression("Histogram of Max " * R[t] * " over [0,10] (δ = 1)"), 
     xlab = expression("Max " * R[t]))

mean(max_Rt)
legend("topright", legend = "mean = 3.812538")
```

### (c)
```{r}
set.seed(1)

delta = 1.2
K = 2
T = 10
M <- 100  # Number of sample paths
t <- seq(0, T, by = dt)

Rt_matrix <- sapply(1:M, function(m) Rt_sim(delta, N, dt)) # row: t, col: mth path of Rt

matplot(t, Rt_matrix, type = "l", lty = 1, col = rainbow(M),
        main = "Sample Paths of Bessel Process (δ = 1.2)",
        xlab = "Time t", ylab = expression(R[t]))

tk_index <- (1:10)/(10/1000) + 1
Rtk <- Rt_matrix[tk_index, ]  # rows: tk, cols: paths

tk_payoff <- pmax(Rtk - K, 0)  # (Rt - K)+

exercise_time <- apply(tk_payoff, 2, function(payoffs) {
  idx <- which(payoffs > 0)[1]  # first positive payoff
  if (is.na(idx)) return(NA)   # never exercised
  return(idx)  # tk index (1 to 10)
})

# Convert index to actual time τ
tau <- ifelse(is.na(exercise_time), NA, exercise_time)  # τ in 1:10

hist(tau, prob = T, breaks = 20)
lines(density(tau[!is.na(tau)]))
summary(tau) # median = 4, mean = 3.967
```

### (d)
```{r}
# Bermuda call option price
Bermuda_payoffs <- sapply(1:M, function(m) {
  payoffs <- tk_payoff[, m]
  holding_values <- numeric(10)
  holding_values[10] <- payoffs[10]

  for (t in 9:1) {
    if (payoffs[t] > 0) {
      immediate_exercise <- payoffs[t]
      future_payoff <- holding_values[t + 1]
      holding_values[t] <- max(immediate_exercise, future_payoff)
    } else {
      holding_values[t] <- holding_values[t + 1]
    }
  }
  return(holding_values[1])
})

Bermuda_price <- mean(Bermuda_payoffs)
cat("Bermuda call price =", Bermuda_price, "\n")

# European call option price = E((R10-K)+)
European_payoff <- tk_payoff[10,]
European_price <- mean(European_payoff)
cat("European call price =", European_price, "\n")


diff_price = Bermuda_price - European_price
cat("Extra cost of Bermuda over European =", diff_price, "\n")

```

###（d)
```{r}
# Bermuda option with optimal exercise
bermuda_payoffs <- numeric(M)
for (m in 1:M) {
  payoffs <- tk_payoff[,m]
  
  # At maturity, optimal decision is to exercise if in-the-money
  continuation_values <- numeric(10)
  continuation_values[10] <- payoffs[10]
  
  for (t in 9:1) {
    # Estimate continuation value at time t
    in_money_paths <- payoffs[t] > 0
    if (sum(in_money_paths) > 0) {
      # For paths that are in-the-money at time t, compare immediate exercise vs. continuation
      immediate_exercise <- payoffs[t]
      # Simple estimate of continuation value: average of future payoffs
      future_payoff <- continuation_values[t+1]
      continuation_values[t] <- max(immediate_exercise, future_payoff)
    } else {
      continuation_values[t] <- continuation_values[t+1]
    }
  }
  
  # Record the payoff from optimal exercise policy
  bermuda_payoffs[m] <- continuation_values[1]
}

bermuda_price <- mean(bermuda_payoffs)
bermuda_price
```

```{r}
#European option payoff: just (R_{10} - K)+ at final time t=10 for each path
# Because your Rt_matrix is (N+1) rows from t=0..10, 
# So, we use the final row is Rt_matrix[nrow(Rt_matrix), ]:
tk_payoff_10 <- tk_payoff[10,]
mean(tk_payoff_10)
```

```{r} 
#Bermuda price:
cat("Bermuda call price =", bermuda_price, "\n")

#European price:
european_price <- mean(tk_payoff_10)
cat("European call price =", european_price, "\n")

#Difference（How much more expensive Bermuda Price than an European price)
premium <- bermuda_price - european_price
cat("Extra cost of Bermuda over European =", premium, "\n")
```

### (e)

#shortcoming of algorithms:

(1)Limited Number of Paths (Monte Carlo Error)
If number of paths is relatively small (e.g. we use M=100 in this case), the variance of the estimator might be high.
Checking for confidence intervals or repeating with bigger M can highlight whether results have converged.

(2)no proper modeling of expected future value
We used a single “average of future payoffs” or just “the next time step’s payoff,” that might be a coarse approximation.
The more accurate method in continuous models is to do a regression across all paths’ next‐step (or multiple‐step) payoffs on the underlying Rt.

(3)Rt_sim

(4)variance reduction

```{r}
set.seed(1)

Rt_sim_stratified <- function(delta, N, T=10) {
  dt <- T / N
  u_vec <- numeric(N)
  for(i in 1:N) {
    lower <- (i-1)/N
    upper <- i/N
    u_vec[i] <- runif(1, min=lower, max=upper)
  }
  z_vec <- qnorm(u_vec)
  dW_vec <- sqrt(dt) * z_vec
  Y <- numeric(N+1)
  Y[1] <- 1
  for(i in 1:N) {
    Y[i+1] <- max(Y[i] + delta*dt + 2*sqrt(Y[i])*dW_vec[i], 0)
  }
  return(sqrt(Y))
}
#Stratified sampling approach
Rt_matrix_strat <- sapply(1:M, function(m) Rt_sim_stratified(delta, N, T))
Rtk_strat <- Rt_matrix_strat[tk_index, ]
tk_payoff_strat <- pmax(Rtk_strat - K, 0)

bermuda_payoffs_strat <- numeric(M)
for (m in 1:M) {
  payoffs <- tk_payoff_strat[,m]
  cont_val <- numeric(10)
  cont_val[10] <- payoffs[10]
  for(t in 9:1) {
    if (payoffs[t]>0) {
      immediate <- payoffs[t]
      future    <- cont_val[t+1]
      cont_val[t] <- max(immediate, future)
    } else {
      cont_val[t] <- cont_val[t+1]
    }
  }
  bermuda_payoffs_strat[m] <- cont_val[1]
}
bermuda_price_strat <- mean(bermuda_payoffs_strat)

#compare sample variance:
cat("Var(Original_Variance) =", var(bermuda_payoffs),
    "Var(Stratified_Variance) =", var(bermuda_payoffs_strat), "\n")
```

#Importance Sampling








